### 2019-11-15 index.Rmd
initial repo setup

---
title: "SI: Analysis Scripts Index"
author: "ADD IN AUTHORS"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    css: stylesheet.css
    fig_caption: yes
    number_sections: no
    toc: no
editor_options:
  chunk_output_type: inline
---

<a href="https://github.com/ORGANIZATION/REPOSITORY/"><img style="position: absolute; top: 0; right: 0; border: 0;" src="https://camo.githubusercontent.com/38ef81f8aca64bb9a64448d0d70f1308ef5341ab/68747470733a2f2f73332e616d617a6f6e6177732e636f6d2f6769746875622f726962626f6e732f666f726b6d655f72696768745f6461726b626c75655f3132313632312e706e67" alt="Fork me on GitHub" data-canonical-src="https://s3.amazonaws.com/github/ribbons/forkme_right_darkblue_121621.png"></a>

# - [analysis 1](analysis1.html)

# - [analysis 2](analysis2.html)

# - [analysis 3](analysis3.html)