### 2019-11-20 isotope_data_reduction.rmd
data and inital data reduction pipeline

---
title: "Isotope data reduction"
date: "`r Sys.Date()`"
  html_document:
    code_folding: show
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
# global knitting options for code rendering
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf"),
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.keep = "all",
  fig.path = file.path("fig_output", paste0(gsub("\\.[Rr]md", "", knitr::current_input()), "_"))
)
```

# Load packages

```{r, message=FALSE, warning=FALSE}
library(tidyverse) # general data wrangling and plotting
library(isoreader) # reading the raw data files
library(isoprocessor) # processing the data
```

This analysis was run using [isoreader](http://isoreader.kopflab.org) version `r packageVersion("isoreader")` and [isoprocessor](http://isoprocessor.kopflab.org/) version `r packageVersion("isoprocessor")`.

All code chunks that contain a critical step towards the final data (i.e. do more than visualization or a data summary) are marked with `(*)` in the header to make it easier to follow all key steps during interactive use.

# Load data

## Read raw data files (*)

```{r, warning=FALSE}
# set file path(s) to data files, folders or rds collections 
# can be multiple folders or mix of folders and files, using example data set here
# (this is a reduced data set of mostly standards with just a few example samples)
data_path <- file.path("data", "201911_isotope_data.cf.rds")

# read files
iso_files_raw <- 
  # path to data files
  data_path %>% 
  # read data files in parallel for fast read
  iso_read_continuous_flow(parallel = TRUE) %>%
  # filter out files with read errors (e.g. from aborted analysis)
  iso_filter_files_with_problems()
```