### 2019-11-20 isotope_data_reduction_TEMPLATE.Rmd

---
title: "Isotope data reduction"
date: "`r Sys.Date()`"
  html_document:
    code_folding: show
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
# global knitting options for code rendering
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf"),
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.keep = "all",
  fig.path = file.path("fig_output", paste0(gsub("\\.[Rr]md", "", knitr::current_input()), "_"))
)
```

# Load packages

```{r, message=FALSE, warning=FALSE}
library(tidyverse) # general data wrangling and plotting
library(isoreader) # reading the raw data files
library(isoprocessor) # processing the data
```

This analysis was run using [isoreader](http://isoreader.kopflab.org) version `r packageVersion("isoreader")` and [isoprocessor](http://isoprocessor.kopflab.org/) version `r packageVersion("isoprocessor")`.

For use as a data processing template, please follow the `Source` link above, download the raw file and adapt as needed. Knitting for stand-alone data analysis works best to `HTML` rather than the in-package default `html_vignette`.

All code chunks that contain a critical step towards the final data (i.e. do more than visualization or a data summary) are marked with `(*)` in the header to make it easier to follow all key steps during interactive use.

# Load data

## Read raw data files (*)

```{r, warning=FALSE}
# set file path(s) to data files, folders or rds collections 
# can be multiple folders or mix of folders and files, using example data set here
# (this is a reduced data set of mostly standards with just a few example samples)
data_path <- iso_get_processor_example("gc_irms_example_carbon.cf.rds")

# read files
iso_files_raw <- 
  # path to data files
  data_path %>% 
  # read data files in parallel for fast read
  iso_read_continuous_flow(parallel = TRUE) %>%
  # filter out files with read errors (e.g. from aborted analysis)
  iso_filter_files_with_problems()
```

## Process file info & peak table (*)

```{r}
# process file information
iso_files <- iso_files_raw %>% 
  # rename key file info columns
  iso_rename_file_info(id1 = `Identifier 1`, id2 = `Identifier 2`) %>% 
  # parse text info into numbers
  iso_parse_file_info(number = Analysis) %>% 
  # process other file information that is specific to the naming conventions
  # of this particular sequence
  iso_mutate_file_info(
    # what is the type of each analysis?
    type = case_when(
      str_detect(id1, "[Zz]ero")      ~ "on_off",
      str_detect(id1, "[Ll]inearity") ~ "lin",
      str_detect(id1, "A5")           ~ "std",
      TRUE                            ~ "sample"
    ),
    # was there seed oxidation?
    seed_oxidation = ifelse(`Seed Oxidation` == "1", "yes", "no"),
    # what was the injection volume based on the AS method name?
    injection_volume = str_extract(`AS Method`, "AS PTV [0-9.]+") %>% 
      parse_number() %>% iso_double_with_units("uL"),
    # what was the concentration? (assuming Preparation = concentration or volume)
    concentration = str_extract(Preparation, "[0-9.]+ ?ng( per |/)uL") %>% 
      parse_number() %>% iso_double_with_units("ng/uL"),
    # or the volume?
    volume = str_extract(Preparation, "[0-9.]+ ?uL") %>% 
      parse_number() %>% iso_double_with_units("uL"),
    # what folder are the data files in? (assuming folder = sequence)
    folder = basename(dirname(file_path))
  ) %>% 
  # add in additional sample metadata (could be any info)
  # note: this would typically be stored in / read from a csv or excel file
  iso_add_file_info(
    tibble::tribble(
      # column names
      ~id1,                           ~sample,          ~fraction,
      # metadata (row-by-row)
      "OG268_CZO-SJER_F1",            "SJER",           "F1",
      "OG271_CZO-CTNA_F1",            "CTNA",           "F1",
      "OG281_CZO-Niwot_Tundra2_F1",   "Niwot_Tundra2",  "F1",
      "OG282_CZO-Niwot_Tundra1_F1",   "Niwot_Tundra1",  "F1"
    ),
    join_by = "id1"
  ) %>% 
  # focus only on the relevant file info, discarding the rest
  iso_select_file_info(
    folder, Analysis, file_datetime, id1, id2, type, sample, 
    seed_oxidation, injection_volume, concentration, volume
  )

# set peak table from vendor data table with default isodat template
iso_files <- iso_set_peak_table_from_isodat_vendor_data_table(iso_files) %>% 
  # convert units from mV to V for amplitudes and area
  iso_convert_peak_table_units(V = mV, Vs = mVs)
```

## Show file information

```{r}
# display file information
iso_files %>% 
  iso_get_file_info() %>% select(-file_id, -folder) %>% 
  iso_make_units_explicit() %>% knitr::kable()
```

## Example chromatograms

```{r "example_chromatograms", fig.width=8, fig.height=8}
# plot the chromatograms
iso_plot_continuous_flow_data(
  # select a few analyses (these #s must exist!)
  iso_filter_files(iso_files, Analysis %in% c(2046, 2106, 2107)),
  # select data and aesthetics
  data = c(44), color = id1, panel = Analysis,
  # zoom in on time interval
  time_interval = c(1000, 3200),
  # peak labels for all peaks > 2V
  peak_label = iso_format(rt, d13C, signif = 3),
  peak_label_size = 3,
  peak_label_filter = amp44 > 2000
) +
  # customize resulting ggplot
  theme(legend.position = "bottom")
```

# ON/OFFs

```{r "on_off_chromatograms", fig.width=8, fig.height=4}
# find files with zero in the Identifier 1 field
on_offs <- iso_files %>% iso_filter_files(type == "on_off")

# visualize the on/offs  
iso_plot_continuous_flow_data(on_offs, data = 44, color = id2)
```

## Summary

```{r "on_off_summary", fig.width=6, fig.height=4}
# calculate on/off summary
on_off_summary <- 
  on_offs %>% 
  # retrieve peak table
  iso_get_peak_table(
    select = c(amp44, area44, d13C),
    include_file_info = c(file_datetime, id2)
  ) %>% 
  # summarize information
  group_by(file_datetime, id2) %>% 
  iso_summarize_data_table()

# table
on_off_summary %>% iso_make_units_explicit() %>% knitr::kable(digits = 3)

# plot
on_off_summary %>% 
  # use generic data plot function
  iso_plot_data(
    x = file_datetime, y = `d13C sd`,
    size = `amp44 mean`, color = id2,
    points = TRUE,
    date_labels = "%b %d - %H:%M"
  ) + 
  # customize resulting ggplot
  expand_limits(y = 0) 
```

# Linearity

```{r "linearity_chromatograms", fig.width=8, fig.height=4}
# find files with linearity in the Identifier 1 field
lins <- iso_files %>% iso_filter_files(type == "lin")

# visualize the linearity runs
iso_plot_continuous_flow_data(lins, data = 44, color = format(file_datetime))
```

## Summary

```{r "linearity_summary", fig.width=7, fig.height=5}
# calculate linearity summary
linearity_summary <- 
  lins %>% 
  # retrieve data table
  iso_get_peak_table(
    select = c(amp44, area44, d13C),
    include_file_info = c(file_datetime)
  ) %>% 
  # calculate linearity regression lines
  nest(data = -file_datetime) %>% 
  mutate(
    fit = map(data, ~lm(d13C ~ amp44, data = iso_strip_units(.x))),
    coefs = map(fit, broom::tidy)
  )

# table
linearity_summary %>%
  unnest(coefs) %>%
  select(file_datetime, term, estimate, std.error) %>%
  filter(term == "amp44") %>%
  mutate(
    term = "linearity [permil/V]",
    estimate = signif(estimate, 2),
    std.error = signif(std.error, 1)
  ) %>% 
  knitr::kable(digits = 4)

# plot
linearity_summary %>% 
  unnest(data) %>% 
  # use generic data plot function
  iso_plot_data(
    x = amp44, y = d13C,
    color = format(file_datetime),
    points = TRUE,
    panel = file_datetime,
    geom_smooth(method = "lm")
  )
```

# Samples and Standards

## Load peak maps

```{r}
# this information is often maintained in a csv or Excel file instead
# but generated here from scratch for demonstration purposes
peak_maps <- 
  tibble::tribble(
    # column names -> defining 2 peak maps, 'std' and 'sample' here
    # but could define an unlimited number of different maps
    ~compound, ~ref_nr, ~`rt:std`,  ~`rt:sample`,
    # peak map data (row-by-row)
    "CO2",			1,		  79,				  79,
    "CO2",			2,	  	178,				178,
    "CO2",			3,	  	278,				278,
    "CO2",			4,	  	377,				377,
    "CO2",			5,	  	477,				477,
    "CO2",			6,	  	576,				576,
    "CO2",			7,	  	676,				676,
    "CO2",			8,	  	775,				775,
    "CO2",			9,	  	875,				875,
    "nC16",			NA,	    1156,				1156,
    "nC17",			NA,    	1280,				1280,
    "nC18",			NA,    	1409,				1409,
    "nC19",			NA,    	1540,				1540,
    "nC20",			NA,    	1670,				1670,
    "nC21",			NA,	    1797,				1797,
    "nC22",			NA,	    1921,				1921,
    "nC23",			NA,	    2040,				2040,
    "nC24",			NA,	    2156,				2156,
    "nC25",			NA,	    2267,				2267,
    "nC26",			NA,	    2375,				2375,
    "nC27",			NA,	    2479,				2479,
    "nC28",			NA,	    2580,				2580,
    "nC29",			NA,	    2674,				2678,
    "nC30",			NA,	    2770,				2773,
    "nC31",			NA,	    NA,				  2863,
    "CO2",			10,	  	3552,				3552,
    "CO2",			11,	  	3592,				3592
  )
peak_maps %>% knitr::kable(digits = 0)
```

## Fetch peak table (*)

```{r}
# identify peaks
peak_table_w_ids <- iso_files %>% 
  # focus on standards and samples only
  iso_filter_files(type %in% c("std", "sample")) %>% 
  # filter out analyses that
  ## a) did not inject
  iso_filter_files(!Analysis %in% c(2031, 2040, 2069, 2098, 2116, 2138)) %>%
  ## b) had double or rogue peaks from mis-injection
  iso_filter_files(!Analysis %in% c(2044, 2132, 2134)) %>% 
  # map peaks
  iso_map_peaks(peak_maps, map_id = type) %>% 
  # peak table
  iso_get_peak_table(include_file_info = everything())
```

## Inspect peak mappings

```{r "example_chromatograms_with_peaks", fig.width = 8, fig.height = 8}
# use same plotting as earlier, except now with peak features added in
iso_files %>% 
  iso_filter_files(Analysis %in% c(2046, 2106, 2107)) %>% 
  iso_plot_continuous_flow_data(
    # select data and aesthetics
    data = 44, color = id1, panel = Analysis,
    # zoom in on time interval
    time_interval = c(1000, 3200),
    # provide our peak table with ids
    peak_table = peak_table_w_ids, 
    # define peak labels, this can be any valid expression
    peak_label = iso_format(id = peak_info, d13C), 
    # specify which labels to show (removing the !is_identified or putting a 
    # restriction by retention time can help a lot with busy chromatograms)
    peak_label_filter = is_identified | !is_identified | is_missing
  )  +
  # customize resulting ggplot
  theme(legend.position = "bottom")
```

## Overview of unclear peak mappings

```{r, eval = FALSE}
# look at missing and unidentified peaks summary (not run during knitting)
# -> check if any of the missing peaks are unexpected (should be there but aren't)
# -> check if any unidentified peaks should be added to the peak maps
# -> check if any ambiguous peaks need clearer peak map retention times
# when in doubt, look at the chromatograms of the problematic files!
# as needed: iterate on peak map update, then peak mapping inspection
peak_table_w_ids %>% 
  iso_get_problematic_peak_mappings() %>% 
  iso_summarize_peak_mappings()
```

# Reference peaks

```{r "ref_peak_variation", fig.width=10, fig.height=4}
# examine variation in the reference peaks
peak_table_w_ids %>% 
  # focus on reference peaks only
  filter(!is.na(ref_nr)) %>% 
  # mark ref peak used for raw delta values (assuming the same in all anlysis)
  mutate(
    ref_info = paste0(ref_nr, ifelse(is_ref == 1, "*", "")) %>% factor() %>% fct_inorder()
  ) %>% 
  # visualize
  iso_plot_ref_peaks(
    # specify the ratios to visualize
    x = Analysis, ratio = `r45/44`, fill = ref_info,
    panel_scales = "fixed"
  ) +
  labs(x = "Analysis", fill = "Reference\npeak")
```

# Analyte peaks

## Select analyte peaks (*)

```{r}
# focus on analyte peaks and calculate useful summary statistics
peak_table_analytes <- peak_table_w_ids %>% 
  # omit reference peaks for further processing (i.e. analyte peaks only)
  filter(is.na(ref_nr)) %>% 
  # for each analysis, calculate a few useful summary statistics
  iso_mutate_peak_table(
    group_by = file_id,
    rel_area = area44 / sum(area44, na.rm = TRUE),
    # calculate mean area for all peaks
    mean_area = mean(area44, na.rm = TRUE),
    # calculate mean area and amplitude for identified peaks
    mean_area_identified = mean(area44[!is.na(compound)], na.rm = TRUE),
    mean_amp_identified = mean(amp44[!is.na(compound)], na.rm = TRUE)
  )
```

## First look

```{r "first_look_sample_and_standard_peaks", fig.width=7, fig.height=6}
peak_table_analytes %>% 
  # focus on identified peaks (comment out this line to see ALL peaks)
  filter(!is.na(compound)) %>% 
  # visualize with convenience function iso_plot_data
  iso_plot_data(
    # choose x and y (multiple y possible)
    x = area44, y = c(amp44, d13C),
    # choose aesthetics
    color = compound, shape = type, label = compound, size = 3,
    # decide what geoms to include
    points = TRUE
  ) +
  # further customize ggplot with a log scale x axis
  scale_x_log10()
```

## Optionally - use interactive plot

```{r, eval=FALSE, fig.width=7, fig.height=6}
# optinally, use an interactive plot to explore your data
# - make sure you install the plotly library --> install.packages("plotly")
# - switch to eval=TRUE in the options of this chunk to include in knit
# - this should work for all plots in this example processing file
library(plotly)
ggplotly(dynamicTicks = TRUE)
```

## Evaluate signal yield

```{r "yield", fig.width=7, fig.height=6}
# evaluate yield
# (only makes sense if both injection volume and concentration are available)
peak_table_analytes %>% 
  # focus on standards
  filter(type == "std") %>% 
  # calculate injection amount
  mutate(injection = iso_double_with_units(
    as.numeric(injection_volume) * as.numeric(concentration), units = "ng")) %>% 
  # visualize
  iso_plot_data(
    # x and y
    x = injection, y = c(mean_area_identified, mean_amp_identified),
    # aesthetics
    color = factor(injection_volume), shape = factor(concentration), 
    size = file_datetime,
    points = TRUE
  ) +
  # modify plot with trendline and axis limits
  geom_smooth(method = "lm", mapping = aes(color = NULL, shape = NULL)) +
  expand_limits(y = 0, x = 0)
```

# Isotope standard values

## Load isotope standards

```{r}
# this information is often maintained in a csv or Excel file instead
# but generated here from scratch for demonstration purposes
standards <- 
  tibble::tribble(
    ~compound,   ~true_d13C,
    "nC16",			-26.15,
    "nC17",			-31.88,
    "nC18",			-32.70,
    "nC19",			-31.99,
    "nC20",			-33.97,
    "nC21",			-28.83,
    "nC22",			-33.77,
    "nC23",			-33.37,
    "nC24",			-32.13,
    "nC25",			-28.46,
    "nC26",			-32.94,
    "nC27",			-30.49,
    "nC28",			-33.20,
    "nC29",			-29.10,
    "nC30",			-29.84
  ) %>% 
  mutate(
    type = "std", 
    true_d13C = iso_double_with_units(true_d13C, "permil")
  )
standards %>% knitr::kable(digits = 2)
```

## Add isotope standards (*)

```{r}
peak_table_w_stds <- 
  peak_table_analytes %>% 
  iso_add_standards(stds = standards, match_by = c("type", "compound")) 
```

# Calibration

## Single analysis calibration (for QA)

### Generate calibrations

Determine calibrations fits for all individual standard analyses.

```{r}
stds_w_calibs <- peak_table_w_stds %>%
  # focus on standards
  filter(type == "std") %>% 
  # remove unassigned peaks
  iso_remove_problematic_peak_mappings() %>% 
  # prepare for calibration by defining the grouping column(s) 
  iso_prepare_for_calibration(group_by = Analysis) %>% 
  # run calibration
  iso_generate_calibration(model = lm(d13C ~ true_d13C)) %>% 
  # unnest some useful file information for visualization
  iso_get_calibration_data(
    select = c(file_datetime, injection_volume, seed_oxidation, mean_area_identified))

# check for problematic calibrations
stds_w_calibs %>% iso_get_problematic_calibrations() -> problematic.calibs
# View(problematic.calibs)

# move forward only with good calibrations
stds_w_calibs <- stds_w_calibs %>% iso_remove_problematic_calibrations()
```

### Coefficients

```{r "single_analysis_calibration_coefficients", fig.width=7, fig.height=8}
# look at coefficients and summary
stds_w_calibs %>% 
  # unnest calibration parameters
  iso_get_calibration_parameters(
    select_from_coefs = 
      c(term, estimate, SE = std.error, signif),
    select_from_summary = 
      c(fit_R2 = adj.r.squared, fit_RMSD = deviance, residual_df = df.residual)) %>%
  arrange(term) %>% 
  knitr::kable(digits = 4)

# visualize the calibration coefficients
stds_w_calibs %>% 
  # plot calibration parameters
  iso_plot_calibration_parameters(
    # x-axis, could also be e.g. Analysis or file_datetime (also use date_breaks!)
    x = mean_area_identified,
    # aesthetics
    color = seed_oxidation, size = mean_area_identified,
    # highlight RSD threshold
    geom_hline(data = tibble(term = factor("RSD"), threshold = 0.5),
               mapping = aes(yintercept = threshold), linetype = 2)
  ) 
```

### Residuals

```{r "single_analysis_residuals", fig.width=7, fig.height=6}
stds_w_calibs %>% 
  # pull out all peak data to including residuals
  iso_get_calibration_data() %>% 
  # focus on standard peaks
  filter(is_std_peak) %>% 
  # calculate parameters to visualize
  iso_mutate_peak_table(
    group_by = Analysis,
     `Var: residual` = resid,
    `Var: area diff from mean` = 
      iso_double_with_units((area44/mean(area44) - 1) * 100, "%")
  ) %>% 
  # visualize
  iso_plot_data(
    # x and y 
    x = compound, y = starts_with("Var"),
    # grouping and aesthetics
    group = paste(Analysis, calib), color = seed_oxidation, linetype = calib,
    # geoms
    lines = TRUE
  )
```

## Global Calibration with all standards

### Generate calibrations (*)

```{r}
# define a global calibration across all standards
global_calibs <- 
  peak_table_w_stds %>%
  # remove (most) problematic peaks to speed up calibration calculations
  # note: could additionally remove unidentified peaks if they are of no interest
  iso_remove_problematic_peak_mappings(remove_unidentified = FALSE) %>% 
  # prepare for calibration (no grouping to go across all analyses)
  iso_prepare_for_calibration() %>% 
  # run different calibrations
  iso_generate_calibration(
    model = c(
      linear = lm(d13C ~ true_d13C),
      with_area = lm(d13C ~ true_d13C + area44),
      with_area_cross = lm(d13C ~ true_d13C * area44)
    ),
    # specify which peaks to include in the calibration, here:
    # - all std_peaks (this filter should always be included!)
    # - standard peaks between 5 and 100 Vs
    # - only analyses with seed oxidation (same as all the samples)
    use_in_calib = is_std_peak & area44 > iso_double_with_units(5, "Vs") & 
      area44 < iso_double_with_units(100, "Vs") & seed_oxidation == "yes"
  ) %>% 
  iso_remove_problematic_calibrations()
```

### Coefficients

```{r "global_calibration_coefficients", fig.width=7, fig.height=12}
# look at coefficients and summary
global_calibs %>% 
  # unnest calibration parameters
  iso_get_calibration_parameters(
    select_from_coefs = 
      c(term, estimate, SE = std.error, signif),
    select_from_summary = 
      c(fit_R2 = adj.r.squared, fit_RMSD = deviance, residual_df = df.residual)) %>%
  arrange(term) %>% 
  knitr::kable(digits = 4)

# visualize coefficients for the different global calibrations
global_calibs %>% iso_plot_calibration_parameters()
```

### Residuals

```{r "global_calibration_residuals", fig.width=8, fig.height=4}
global_calibs %>% 
  iso_plot_residuals(
    x = compound, group = paste(Analysis, calib), 
    points = FALSE, lines = TRUE,
    trendlines = FALSE, value_ranges = FALSE
  ) 
```

### Apply global calibration (*)

```{r, cache=TRUE}
# note that depending on the number of data points, this may take a while
# for faster calculations, chose calculate_error = FALSE
global_calibs_applied <- 
  global_calibs %>% 
  # which calibration to use? can include multiple if desired to see the result
  # in this case, the area conscious calibrations are not necessary
  filter(calib == "linear") %>% 
  # apply calibration indication what should be calcculated
  iso_apply_calibration(true_d13C, calculate_error = TRUE)

# calibration ranges
global_calibs_with_ranges <-
  global_calibs_applied %>% 
  # evaluate calibration range for the measured area.Vs and predicted d13C
  iso_evaluate_calibration_range(area44, true_d13C_pred)

# show calibration ranges
global_calibs_with_ranges %>% 
  iso_get_calibration_range() %>% 
  iso_remove_list_columns() %>% 
  knitr::kable(d = 2)

# create calibrated peak table
peak_table_calibrated <- global_calibs_with_ranges %>% 
  iso_get_calibration_data()
```

# Evaluation

## Overview

```{r "data_overview", fig.width=7, fig.height=9}
# replicate earlier overview plot but now with the calibrated delta values
# and with a higlight of the calibration ranges and which points are in range
peak_table_calibrated %>% 
  # focus on identified peaks (comment out this line to see ALL peaks)
  filter(!is.na(compound)) %>% 
  # visualize with convenience function iso_plot_data
  iso_plot_data(
    # choose x and y (multiple y possible)
    x = area44, y = true_d13C_pred,
    # choose aesthetics
    color = in_range, shape = type, label = compound, size = 3,
    # decide what geoms to include
    points = TRUE
  ) %>% 
  # highlight calibration range
  iso_mark_calibration_range() +
  # further customize ggplot with a log scale x axis
  scale_x_log10() +
  # legend
  theme(legend.position = "bottom", legend.direction = "vertical")
```

## Standards

```{r "standards", fig.width=7, fig.height=6}
# visualize how standard line up after calibration
peak_table_calibrated %>% 
  # focus on peaks in calibration
  filter(in_calib) %>% 
  # visualize
  iso_plot_data(
    # x and y
    x = true_d13C, y = true_d13C_pred, y_error = true_d13C_pred_se,
    # aesthetics
    color = compound, size = area44,
    # geoms
    points = TRUE,
    # add 1:1 trendline
    geom_abline(slope = 1, intercept = 0)
  ) 
```

## Data

### Isotopic values

```{r "samples", fig.width = 10, fig.height = 7}
# Warning: data outside the calibration range MUST be taken with a
# grain of salt, especially at the low signal area/amplitude end
peak_table_calibrated %>% 
  # focus on samples
  filter(type == "sample") %>% 
  # focus on identified peaks (comment out this line to see ALL peaks)
  # if there's a lot of data, might need to hone in on a few at a time
  filter(!is.na(compound)) %>% 
  # visualize
  iso_plot_data(
    # x and y
    x = sample, y = true_d13C_pred, y_error = true_d13C_pred_se,
    # aesthetics
    color = in_range, size = rel_area, 
    # paneling
    panel = compound, panel_scales = "fixed",
    # geoms
    points = TRUE
  ) %>% 
  # mark calibration range (include optionally)
  iso_mark_calibration_range() +
  # color palette (here example of manual: www.google.com/search?q=color+picker)
  scale_color_manual(
    values = c("#984EA3", "#E41A1C", "#E41A1C", "#377EB8", "#4DAF4A")
  ) +
  # clarify size scale
  scale_size_continuous(breaks = c(0.01, 0.05, 0.1, 0.2, 0.3), 
                        labels = function(x) paste0(100*x, "%")) +
  # plot labels
  labs(x = NULL)
```

### Relative abundances

```{r "abundances", fig.width = 7, fig.height = 6, warning=FALSE}
# visualize relative abundances
peak_table_calibrated %>% 
  # focus on samples
  filter(type == "sample") %>% 
  # focus on identified peaks (comment out this line to see ALL peaks)
  #filter(!is.na(compound)) %>% 
  # visualize
  iso_plot_data(
    # x and y
    x = sample, y = rel_area,
    # aesthetics
    fill = compound,
    # barchart
    geom_bar(stat = "identity")
  ) +
  # scales
  scale_y_continuous(labels = function(x) paste0(100*x, "%"), expand = c(0, 0)) +
  # plot labels
  labs(x = NULL, y = "Relative abundance (by area)")
```

### Summary

```{r}
# generate data summary
peak_data <- 
  peak_table_calibrated %>% 
  # focus on identified peaks in the samples
  filter(type == "sample", !is.na(compound)) %>% 
  select(sample, compound, Analysis, 
         area44, true_d13C_pred, true_d13C_pred_se, in_range) %>% 
  arrange(sample, compound, Analysis)

# summarize replicates
# this example data set does not contain any replicates but typically all
# analyses should in which case a statistical data summary can be useful
peak_data_summary <- 
  peak_data %>% 
  # here: only peaks within the area calibration range are included
  # you have to explicitly decide which peaks you trust if they are out of range
  filter(
    in_range %in% c("in range", "<'true_d13C_pred' range", ">'true_d13C_pred' range")
  ) %>% 
  # summarize for each sample and compound
  group_by(sample, compound) %>% 
  iso_summarize_data_table(area44, true_d13C_pred)

peak_data_summary %>% iso_make_units_explicit() %>% knitr::kable(d = 2)
```

# Final

Final data processing and visualization usually depends on the type of data and the metadata available for contextualization (e.g. core depth, source organism, age, etc.). The relevant metadata can be added easily with `iso_add_file_info()` during the initial data load / file info procesing. Alternatively, just call `iso_add_file_info()` again at this later point or use dplyr's `left_join` directly.

```{r "final_data", fig.width = 7, fig.height = 6}
# @user: add final data processing and plot(s)
```

# Export

```{r}
# export the global calibration with all its information and data to Excel
peak_table_calibrated %>% 
  iso_export_calibration_to_excel(
    filepath = format(Sys.Date(), "%Y%m%d_gc_irms_example_carbon_export.xlsx"),
    # include data summary as an additional useful tab
    `data summary` = peak_data_summary
  )
```