### 2019-12-02 abundance.Rmd
secondcommit

---
title: "Analysis: NAME"
subtitle: "SUBTITLE"
author: "AUTHORS"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged
    css: stylesheet.css
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_float: true
    toc_depth: 3
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# load libraries
library(tidyverse) 
library(latex2exp)
library(readxl)
library(isoreader)

# source functions
source(file.path("libs", "visualization.R"))
source(file.path("libs", "chromleon.R"))

# set global chunk options for figure output
figure_prefix <- "PREFIX-"
knitr::opts_chunk$set(
  dev=c("png", "pdf"), dev.args=list(pdf = list(encoding="WinAnsi", useDingbats=FALSE)),
  fig.keep="all", fig.path=file.path("figures", figure_prefix))
```

# Load data

```{r}
root <- "data"
all_data <-
  tibble(
    sample = c("a", "b"),
    file = c("TSQ5934_mAx.xls", "TSQ5934_mAx.xls"),
    data = map(file.path(root, file), read_chromeleon_export),
    injection_details = map(data, "injection_details"),
    integration_results = map(data, "integration_results")
  ) %>% select(-data)
```

## Show injection details

```{r}
all_data %>% select(-integration_results) %>% unnest(injection_details) %>% 
  knitr::kable()
```

# Analyze data

```{r}
results <- all_data %>% select(-injection_details) %>% unnest(integration_results) %>% 
  filter(nchar(`Peak Name`) > 0)

results %>% 
  filter(!str_detect(`Peak Name`, fixed("?"))) %>% 
  ggplot() +
  aes(`Peak Name`, `Relative Area`) +
  geom_bar(stat = "identity") +
  theme_figure(axis_x_rotate = 90) +
  scale_y_continuous(expand = c(0,0)) +
  facet_wrap(~sample)
```

```{r}
results %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = `Peak Name`) +
  geom_bar(stat = "identity")
```

```{r}
results %>% 
  left_join(
    tibble(
      `Peak Name` = c("19:0", "23:0"), 
      group = c("19", "23")
    ), by = "Peak Name"
  ) %>% 
  ggplot() +
  aes(`Retention Time`, `Relative Height`, fill = group, label = `Peak Name`) +
  geom_bar(stat = "identity") +
  facet_grid(sample ~ .) +
  theme_figure() +
  scale_y_continuous(expand = c(0, 0))
```

```{r}
library(plotly)
ggplotly(ggplot2::last_plot(), dynamicTicks = TRUE)
```