### 2019-12-07 example_R/gc_irms_example.Rmd

---
title: "gc_irms_example"
output: html_document
---

# Set Up, Load packages
```{r}
# restart your R session (this command only works in RStudio)
.rs.restartR()

# installs the development tools package if not yet installed
if(!requireNamespace("devtools", quietly = TRUE)) install.packages("devtools")

# installs the newest version of isoreader and isoprocessor
devtools::install_github("isoverse/isoreader")
devtools::install_github("isoverse/isoprocessor")
```

```{r}
library(tidyverse) # general data wrangling and plotting
library(isoreader) # reading the raw data files
library(isoprocessor) # processing the data
```

# Load Data
```{r}
# Set file paths to data files, folders or rds collections
# Can be multiple folders or mix of folders/files
# this is reduced data, mostly standards, just a few samples
data_path <- iso_get_processor_example("gc_irms_example_carbon.cf.rds")
```

# Read files, create "raw" iso files
Here we read the continuous flow files in parallel (improves speed) and remove files with problems, such as aborted runs.
```{r}
iso_files_raw <-
  data_path %>% # path to data files
  iso_read_continuous_flow(parallel = TRUE) %>% # read data files in parallel for fast read
  iso_filter_files_with_problems() # filter out files with read errors, such as from aborted analysis

```

# Process File Info and Peak Table
```{r}
# process file information
iso_files <- iso_files_raw %>% 
  # rename key file info columns
  iso_rename_file_info(id1 = `Identifier 1`, id2 = `Identifier 2`) %>% 
  # parse text info into numbers
  iso_parse_file_info(number = Analysis) %>% 
  # process other file information that is specific to the naming conventions
  # of this particular sequence
  iso_mutate_file_info(
    # what is the type of each analysis?
    type = case_when(
      str_detect(id1, "[Zz]ero")      ~ "on_off",
      str_detect(id1, "[Ll]inearity") ~ "lin",
      str_detect(id1, "A5")           ~ "std",
      TRUE                            ~ "sample"
    ),
    # was there seed oxidation?
    seed_oxidation = ifelse(`Seed Oxidation` == "1", "yes", "no"),
    # what was the injection volume based on the AS method name?
    injection_volume = str_extract(`AS Method`, "AS PTV [0-9.]+") %>% 
      parse_number() %>% iso_double_with_units("uL"),
    # what was the concentration? (assuming Preparation = concentration or volume)
    concentration = str_extract(Preparation, "[0-9.]+ ?ng( per |/)uL") %>% 
      parse_number() %>% iso_double_with_units("ng/uL"),
    # or the volume?
    volume = str_extract(Preparation, "[0-9.]+ ?uL") %>% 
      parse_number() %>% iso_double_with_units("uL"),
    # what folder are the data files in? (assuming folder = sequence)
    folder = basename(dirname(file_path))
  ) %>% 
  # add in additional sample metadata (could be any info)
  # note: this would typically be stored in / read from a csv or excel file
  iso_add_file_info(
    tibble::tribble(
      # column names
      ~id1,                           ~sample,          ~fraction,
      # metadata (row-by-row)
      "OG268_CZO-SJER_F1",            "SJER",           "F1",
      "OG271_CZO-CTNA_F1",            "CTNA",           "F1",
      "OG281_CZO-Niwot_Tundra2_F1",   "Niwot_Tundra2",  "F1",
      "OG282_CZO-Niwot_Tundra1_F1",   "Niwot_Tundra1",  "F1"
    ),
    join_by = "id1"
  ) %>% 
  # focus only on the relevant file info, discarding the rest
  iso_select_file_info(
    folder, Analysis, file_datetime, id1, id2, type, sample, 
    seed_oxidation, injection_volume, concentration, volume
  )
#> Info: renaming 2 file info column(s) wherever they exist across 45 isofile(s):
#>  - for 44 file(s): 'Identifier 1'->'id1', 'Identifier 2'->'id2'
#>  - for 1 file(s): 'Identifier 1'->'id1'
#> Info: parsing 1 file info columns for 45 data file(s):
#>  - to number: 'Analysis'
#> Info: mutating file info for 45 data file(s)
#> Info: adding new file information ('sample', 'fraction') to 45 data file(s), joining by 'id1'...
#>  - 'id1' join: 4/4 new info rows matched 4/45 data files
#> Info: keeping 12 file info column(s) wherever they exist across 45 isofile(s):
#>  - for 45 file(s): 'file_id', 'folder', 'Analysis', 'file_datetime', 'id1', 'id2', 'type', 'sample', 'seed_oxidation', 'injection_volume', 'concentration', 'volume'

# set peak table from vendor data table with default isodat template
iso_files <- iso_set_peak_table_from_isodat_vendor_data_table(iso_files) %>% 
  # convert units from mV to V for amplitudes and area
  iso_convert_peak_table_units(V = mV, Vs = mVs)
#> Info: setting peak table from vendor data table with the following renames:
#> - for 45 file(s): 'Nr.'->'peak_nr', 'Is Ref.?'->'is_ref', 'Start'->'rt_start', 'Rt'->'rt', 'End'->'rt_end', 'Ampl 44'->'amp44', 'Ampl 45'->'amp45', 'Ampl 46'->'amp46', 'BGD 44'->'bgrd44_start', 'BGD 45'->'bgrd45_start', 'BGD 46'->'bgrd46_start', 'BGD 44'->'bgrd44_end', 'BGD 45'->'bgrd45_end', 'BGD 46'->'bgrd46_end', 'rIntensity 44'->'area44', 'rIntensity 45'->'area45', 'rIntensity 46'->'area46', 'rR 45CO2/44CO2'->'r45/44', 'rR 46CO2/44CO2'->'r46/44', 'rd 45CO2/44CO2'->'rd45/44', 'rd 46CO2/44CO2'->'rd46/44', 'd 45CO2/44CO2'->'d45/44', 'd 46CO2/44CO2'->'d46/44', 'd 13C/12C'->'d13C', 'd 18O/16O'->'d18O', 'd 17O/16O'->'d17O', 'AT% 13C/12C'->'at13C', 'AT% 18O/16O'->'at18O'
#> Info: converting peak table column units where applicable from 'mV'->'V' and 'mVs'->'Vs' for columns matching 'everything()'

```

# Display file info
```{r}
# display file information
iso_files %>% 
  iso_get_file_info() %>% select(-file_id, -folder) %>% 
  iso_make_units_explicit() %>% knitr::kable()
#> Info: aggregating file info from 45 data file(s)
```

# Chromatograms
```{r}
# plot the chromatograms
iso_plot_continuous_flow_data(
  # select a few analyses (these #s must exist!)
  iso_filter_files(iso_files, Analysis %in% c(2046, 2106, 2107)),
  # select data and aesthetics
  data = c(44), color = id1, panel = Analysis,
  # zoom in on time interval
  time_interval = c(1000, 3200),
  # peak labels for all peaks > 2V
  peak_label = iso_format(rt, d13C, signif = 3),
  peak_label_size = 3,
  peak_label_filter = amp44 > 2000
) +
  # customize resulting ggplot
  theme(legend.position = "bottom")
#> Info: applying file filter, keeping 3 of 45 files
```

```{r}
# find files with zero in the Identifier 1 field
on_offs <- iso_files %>% iso_filter_files(type == "on_off") 
#> Info: applying file filter, keeping 4 of 45 files

# visualize the on/offs  
iso_plot_continuous_flow_data(on_offs, data = 44, color = id2)
```

```{r}
# calculate on/off summary
on_off_summary <- 
  on_offs %>% 
  # retrieve peak table
  iso_get_peak_table(
    select = c(amp44, area44, d13C),
    include_file_info = c(file_datetime, id2)
  ) %>% 
  # summarize information
  group_by(file_datetime, id2) %>% 
  iso_summarize_data_table() 
#> Info: aggregating peak table from 4 data file(s), including file info 'c(file_datetime, id2)'

# table
on_off_summary %>% iso_make_units_explicit() %>% knitr::kable(digits = 3)
```

```{r}
# plot
on_off_summary %>% 
  # use generic data plot function
  iso_plot_data(
    x = file_datetime, y = `d13C sd`,
    size = `amp44 mean`, color = id2,
    points = TRUE,
    date_labels = "%b %d - %H:%M"
  ) + 
  # customize resulting ggplot
  expand_limits(y = 0) 
```

```{r}
# find files with linearity in the Identifier 1 field
lins <- iso_files %>% iso_filter_files(type == "lin") 
#> Info: applying file filter, keeping 1 of 45 files

# visualize the linearity runs
iso_plot_continuous_flow_data(lins, data = 44, color = format(file_datetime))
```

```{r}
# calculate linearity summary
linearity_summary <- 
  lins %>% 
  # retrieve data table
  iso_get_peak_table(
    select = c(amp44, area44, d13C),
    include_file_info = c(file_datetime)
  ) %>% 
  # calculate linearity regression lines
  nest(data = -file_datetime) %>% 
  mutate(
    fit = map(data, ~lm(d13C ~ amp44, data = iso_strip_units(.x))),
    coefs = map(fit, broom::tidy)
  )
#> Info: aggregating peak table from 1 data file(s), including file info 'c(file_datetime)'

# table
linearity_summary %>%
  unnest(coefs) %>%
  select(file_datetime, term, estimate, std.error) %>%
  filter(term == "amp44") %>%
  mutate(
    term = "linearity [permil/V]",
    estimate = signif(estimate, 2),
    std.error = signif(std.error, 1)
  ) %>% 
  knitr::kable(digits = 4)
```

```{r}
# plot
linearity_summary %>% 
  unnest(data) %>% 
  # use generic data plot function
  iso_plot_data(
    x = amp44, y = d13C,
    color = format(file_datetime),
    points = TRUE,
    panel = file_datetime,
    geom_smooth(method = "lm")
  )
```

```{r}
# this information is often maintained in a csv or Excel file instead
# but generated here from scratch for demonstration purposes
peak_maps <- 
  tibble::tribble(
    # column names -> defining 2 peak maps, 'std' and 'sample' here
    # but could define an unlimited number of different maps
    ~compound, ~ref_nr, ~`rt:std`,  ~`rt:sample`,
    # peak map data (row-by-row)
    "CO2",          1,        79,                 79,
    "CO2",          2,      178,                178,
    "CO2",          3,      278,                278,
    "CO2",          4,      377,                377,
    "CO2",          5,      477,                477,
    "CO2",          6,      576,                576,
    "CO2",          7,      676,                676,
    "CO2",          8,      775,                775,
    "CO2",          9,      875,                875,
    "nC16",         NA,     1156,               1156,
    "nC17",         NA,     1280,               1280,
    "nC18",         NA,     1409,               1409,
    "nC19",         NA,     1540,               1540,
    "nC20",         NA,     1670,               1670,
    "nC21",         NA,     1797,               1797,
    "nC22",         NA,     1921,               1921,
    "nC23",         NA,     2040,               2040,
    "nC24",         NA,     2156,               2156,
    "nC25",         NA,     2267,               2267,
    "nC26",         NA,     2375,               2375,
    "nC27",         NA,     2479,               2479,
    "nC28",         NA,     2580,               2580,
    "nC29",         NA,     2674,               2678,
    "nC30",         NA,     2770,               2773,
    "nC31",         NA,     NA,               2863,
    "CO2",          10,     3552,               3552,
    "CO2",          11,     3592,               3592
  )
peak_maps %>% knitr::kable(digits = 0)
```

```{r}
# identify peaks
peak_table_w_ids <- iso_files %>% 
  # focus on standards and samples only
  iso_filter_files(type %in% c("std", "sample")) %>% 
  # filter out analyses that
  ## a) did not inject
  iso_filter_files(!Analysis %in% c(2031, 2040, 2069, 2098, 2116, 2138)) %>%
  ## b) had double or rogue peaks from mis-injection
  iso_filter_files(!Analysis %in% c(2044, 2132, 2134)) %>% 
  # map peaks
  iso_map_peaks(peak_maps, map_id = type) %>% 
  # peak table
  iso_get_peak_table(include_file_info = everything())
```

```{r}
# use same plotting as earlier, except now with peak features added in
iso_files %>% 
  iso_filter_files(Analysis %in% c(2046, 2106, 2107)) %>% 
  iso_plot_continuous_flow_data(
    # select data and aesthetics
    data = 44, color = id1, panel = Analysis,
    # zoom in on time interval
    time_interval = c(1000, 3200),
    # provide our peak table with ids
    peak_table = peak_table_w_ids, 
    # define peak labels, this can be any valid expression
    peak_label = iso_format(id = peak_info, d13C), 
    # specify which labels to show (removing the !is_identified or putting a 
    # restriction by retention time can help a lot with busy chromatograms)
    peak_label_filter = is_identified | !is_identified | is_missing
  )  +
  # customize resulting ggplot
  theme(legend.position = "bottom")
#> Info: applying file filter, keeping 3 of 45 files
```

```{r}
# look at missing and unidentified peaks summary (not run during knitting)
# -> check if any of the missing peaks are unexpected (should be there but aren't)
# -> check if any unidentified peaks should be added to the peak maps
# -> check if any ambiguous peaks need clearer peak map retention times
# when in doubt, look at the chromatograms of the problematic files!
# as needed: iterate on peak map update, then peak mapping inspection
peak_table_w_ids %>% 
  iso_get_problematic_peak_mappings() %>% 
  iso_summarize_peak_mappings()
```

```{r}
# examine variation in the reference peaks
peak_table_w_ids %>% 
  # focus on reference peaks only
  filter(!is.na(ref_nr)) %>% 
  # mark ref peak used for raw delta values (assuming the same in all anlysis)
  mutate(
    ref_info = paste0(ref_nr, ifelse(is_ref == 1, "*", "")) %>% factor() %>% fct_inorder()
  ) %>% 
  # visualize
  iso_plot_ref_peaks(
    # specify the ratios to visualize
    x = Analysis, ratio = `r45/44`, fill = ref_info,
    panel_scales = "fixed"
  ) +
  labs(x = "Analysis", fill = "Reference\npeak")
```

# Analyte Peaks
```{r}
# focus on analyte peaks and calculate useful summary statistics
peak_table_analytes <- peak_table_w_ids %>% 
  # omit reference peaks for further processing (i.e. analyte peaks only)
  filter(is.na(ref_nr)) %>% 
  # for each analysis, calculate a few useful summary statistics
  iso_mutate_peak_table(
    group_by = file_id,
    rel_area = area44 / sum(area44, na.rm = TRUE),
    # calculate mean area for all peaks
    mean_area = mean(area44, na.rm = TRUE),
    # calculate mean area and amplitude for identified peaks
    mean_area_identified = mean(area44[!is.na(compound)], na.rm = TRUE),
    mean_amp_identified = mean(amp44[!is.na(compound)], na.rm = TRUE)
  )
```

```{r}
peak_table_analytes %>% 
  # focus on identified peaks (comment out this line to see ALL peaks)
  filter(!is.na(compound)) %>% 
  # visualize with convenience function iso_plot_data
  iso_plot_data(
    # choose x and y (multiple y possible)
    x = area44, y = c(amp44, d13C),
    # choose aesthetics
    color = compound, shape = type, label = compound, size = 3,
    # decide what geoms to include
    points = TRUE
  ) +
  # further customize ggplot with a log scale x axis
  scale_x_log10()
```