### 2019-12-07 isotope_data_reduction.rmd
added new plots and map peaks

---
title: "Isotope data reduction"
date: "`r Sys.Date()`"
  html_document:
    code_folding: show
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
# global knitting options for code rendering
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf"),
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.keep = "all",
  fig.path = file.path("fig_output", paste0(gsub("\\.[Rr]md", "", knitr::current_input()), "_"))
)
```

# Load packages

```{r, message=FALSE, warning=FALSE}
library(tidyverse) # general data wrangling and plotting
library(isoreader) # reading the raw data files
library(isoprocessor) # processing the data
```

This analysis was run using [isoreader](http://isoreader.kopflab.org) version `r packageVersion("isoreader")` and [isoprocessor](http://isoprocessor.kopflab.org/) version `r packageVersion("isoprocessor")`.

All code chunks that contain a critical step towards the final data (i.e. do more than visualization or a data summary) are marked with `(*)` in the header to make it easier to follow all key steps during interactive use.

# Load data

## Read raw data files (*)

```{r, warning=FALSE}
# set file path(s) to data files, folders or rds collections 
# can be multiple folders or mix of folders and files, using example data set here
# (this is a reduced data set of mostly standards with just a few example samples)
data_path <- file.path("data", "201911_isotope_data.cf.rds")

# read files
iso_files_raw <- 
  # path to data files
  data_path %>% 
  # read data files in parallel for fast read
  iso_read_continuous_flow() %>%
  # filter out files with read errors (e.g. from aborted analysis)
  iso_filter_files_with_problems()
```

## Process file info & peak table (*)

```{r}
# process file information
iso_files <- iso_files_raw %>% 
  # rename key file info columns
  iso_rename_file_info(analysis = Analysis, id1 = `Identifier 1`, id2 = `Identifier 2`) %>% 
  # parse text info into numbers
  iso_parse_file_info(number = analysis) %>% 
  # process other file information that is specific to the naming conventions
  # of this particular sequence
  iso_mutate_file_info(
    # what is the type of each analysis?
    type = case_when(
      str_detect(id1, "[Zz]ero")      ~ "on_off",
      str_detect(id1, "H3")           ~ "H3_factor",
      str_detect(id1, "F8")           ~ "std",
      TRUE                            ~ "sample"
    ),
    # what was the concentration? (assuming Preparation = concentration or volume)
    concentration = 
      ifelse(type == "std",  
             str_extract(Preparation, "[0-9.]+ ?ng( per |/)uL") %>% 
               parse_number() %>% iso_double_with_units("ng/uL"),
             NA),
    # what folder are the data files in? (assuming folder = sequence)
    folder = basename(dirname(file_path))
  ) %>% 
  # focus only on the relevant file info, discarding the rest
  iso_select_file_info(
    folder, analysis, file_datetime, id1, type, concentration
  ) %>% 
  # add in additional sample metadata (could be any info)
  # note: this would typically be stored in / read from a csv or excel file
  iso_add_file_info(
    readxl::read_excel(file.path("data", "isotopes_metadata.xlsx")),
    join_by = "id1"
  )

# set peak table from vendor data table with default isodat template
iso_files <- iso_set_peak_table_from_isodat_vendor_data_table(iso_files) %>% 
  # convert units from mV to V for amplitudes and area
  iso_convert_peak_table_units(V = mV, Vs = mVs)

# focus on sample files
sample_files <- iso_filter_files(iso_files, type == "sample" & analysis > 5651)
```

## Show file information

```{r}
# display file information
iso_files %>% 
  iso_get_file_info() %>% select(-file_id, -folder) %>% 
  iso_make_units_explicit() %>% knitr::kable()
```

## Example chromatograms

```{r "example_chromatograms", fig.width=8, fig.height=8}
# plot the chromatograms
iso_plot_continuous_flow_data(
  # all sample files
  sample_files,
  # select data and aesthetics
  data = c(2), color = id1, panel = analysis,
  # zoom in on time interval
  time_interval = c(1000, 3000),
  # peak labels for all peaks > 2V
  peak_bounds = TRUE,
  peak_marker = FALSE,
  peak_label = iso_format(rt),
  #peak_label_size = 3,
  peak_label_filter = analysis == 5685 & amp2 > 1
) 
```

# Peak Mapping

```{r}
# this information is often maintained in a csv or Excel file instead
# but generated here from scratch for demonstration purposes
peak_maps <- 
  readxl::read_excel(file.path("data","isotopes_peak_maps.xlsx"))
peak_maps %>% knitr::kable(digits = 0)

# identify peaks
sample_files_w_ids <- sample_files  %>% 
  # map peaks
  iso_map_peaks(peak_maps, map_id = type)
```

## Chromatograms

```{r "sample_chromatograms", fig.width=8, fig.height=8}
# plot the chromatograms
iso_plot_continuous_flow_data(
  # all sample files
  sample_files_w_ids,
  # select data and aesthetics
  data = c(2), color = id1, panel = analysis,
  # zoom in on time interval
  time_interval = c(1000, 3000),
  # peak labels for all peaks > 2V
  peak_bounds = TRUE,
  peak_marker = FALSE,
  peak_label = iso_format(compound, d2H),
  #peak_label_size = 3,
  peak_label_filter = compound != "unknown"
) 
```

## Data

```{r}
real_data <- 
  sample_files_w_ids %>% 
  iso_get_peak_table(include_file_info = everything()) %>% 
  filter(compound != "H2") %>% 
  group_by(file_id) %>% 
  mutate(rel_amount = area2 / sum(area2[compound != "IS"])) %>% 
  ungroup() %>% 
  filter(compound != "unknown") 
```

```{r}
real_data %>% 
  mutate(sample_label = iso_format(control, inc_time, with_glu)) %>% 
  iso_plot_data(
    x = inc_time, y = d2H, color = compound, shape = with_glu, size = rel_amount,
    points = TRUE,
    lines = TRUE,
    panel = with_glu ~ .
    #panel_scales = "fixed"
  )
```

# FOR PROCESSING ONLY

# Export data

```{r}
irms_example_data <- 
  sample_files_w_ids %>%
  iso_filter_files(analysis == 5689) %>% 
  iso_get_peak_table() %>% 
  filter(compound != "H2") %>% 
  select(rt, rt_start, rt_end, amp2, area2) %>% 
  mutate(rel_height = amp2/sum(amp2) * 100) %>% 
  filter(rel_height > 1)

irms_example_data %>% 
  openxlsx::write.xlsx(file.path("tables", "gc_irms_single_sample_peak_table.xlsx"))
```

# Peak Mapping help

```{r}
tsq_example_data <- read_excel(file.path("tables", "tsq_single_sample_peak_table.xlsx"))
```

```{r}
bind_rows(
  irms_example_data %>% mutate(type = "irms") %>% iso_strip_units(),
  tsq_example_data %>% mutate(type = "tsq")
) %>% 
  ggplot() +
  aes(rt, rel_height, label = paste0(`Peak Name`, "\n", round(rt))) +
  geom_bar(stat = "identity")+
  ggrepel::geom_text_repel(color = "red", alpha = 0.5, min.segment.length = 0)+
  facet_grid(type ~ .)
ggsave(file.path("figures", "irms_tsq_comparison.pdf"), width = 12, height = 8)
```

##
```{r}
# examine variation in the reference peaks
sample_files_w_ids %>% 
  # focus on reference peaks only
  filter(!is.na(ref_nr)) %>% 
  # mark ref peak used for raw delta values (assuming the same in all anlysis)
  mutate(
    ref_info = paste0(ref_nr, ifelse(is_ref == 1, "*", "")) %>% factor() %>% fct_inorder()
  ) %>% 
  # visualize
  iso_plot_ref_peaks(
    # specify the ratios to visualize
    x = Analysis, ratio = `r45/44`, fill = ref_info,
    panel_scales = "fixed"
  ) +
  labs(x = "Analysis", fill = "Reference\npeak")
```

##
```{r}
# focus on analyte peaks and calculate useful summary statistics
peak_table_analytes <- sample_files_w_ids %>% 
  # omit reference peaks for further processing (i.e. analyte peaks only)
  filter(is.na(ref_nr)) %>% 
  # for each analysis, calculate a few useful summary statistics
  iso_mutate_peak_table(
    group_by = file_id,
    rel_area = area44 / sum(area44, na.rm = TRUE),
    # calculate mean area for all peaks
    mean_area = mean(area44, na.rm = TRUE),
    # calculate mean area and amplitude for identified peaks
    mean_area_identified = mean(area44[!is.na(compound)], na.rm = TRUE),
    mean_amp_identified = mean(amp44[!is.na(compound)], na.rm = TRUE)
  )
```

##
```{r}
# find files with linearity in the Identifier 1 field
lins <- iso_files %>% iso_filter_files(type == "lin") 
#> Info: applying file filter, keeping 1 of 45 files

# visualize the linearity runs
iso_plot_continuous_flow_data(lins, data = 44, color = format(file_datetime))

```

##
```{r}
# calculate on/off summary
on_off_summary <- 
  on_offs %>% 
  # retrieve peak table
  iso_get_peak_table(
    select = c(amp44, area44, d13C),
    include_file_info = c(file_datetime, id2)
  ) %>% 
  # summarize information
  group_by(file_datetime, id2) %>% 
  iso_summarize_data_table() 
#> Info: aggregating peak table from 4 data file(s), including file info 'c(file_datetime, id2)'

# table
on_off_summary %>% iso_make_units_explicit() %>% knitr::kable(digits = 3)
```