### 2021-02-17 gc_column_comparison.Rmd
added comparison of gc column elution quality

---
title: "Column Comparison"
author: "Tristan Caro"
date: "2/16/2021"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

```{r}
# Libraries
library(tidyverse)
library(plotly)
library(ggrepel)

# Source
source(file.path("libs", "visualization.R"))
source(file.path("libs", "chromleon.R"))
source(file.path("libs", "chromeleon_peaks.R"))
```

```{r}
# Define path and sample names
root <- "data/2021.02.04_TC_FAMES_Test_DB1"
path_names <- list.files(root)
path_names <- path_names %>% 
  str_subset("blank", negate = TRUE) %>% 
  str_subset("Blank", negate = TRUE) %>% 
  str_subset("GCMS-STD", negate = TRUE)
sample_names <- path_names %>% 
  str_replace("TSQ.....", "") %>% 
  str_replace(".xls", "")

# Create a tibble of integration data
# read_chromeleon_export is a sourced function. See above.
all_data <-
  tibble(
    sample = sample_names,
    file = path_names,
    data = map(file.path(root, file), read_chromeleon_export),
    injection_details = map(data, "injection_details"),
    integration_results = map(data, "integration_results")
  ) %>% select(-data)

# Create a tibble of peak analysis data
peak_data <-
  tibble(
    sample = sample_names,
    file = path_names,
    data = map(file.path(root, file), read_chromeleon_export_peaks),
    injection_details = map(data, "injection_details"),
    peak_analysis = map(data, "peak_analysis")
  ) %>% select(-data)
```

```{r, warning = FALSE}
# Data cleanup
results <- all_data %>% 
  select(-injection_details) %>% 
  unnest(integration_results) %>% 
  filter(nchar(`Peak Name`) > 0) %>% 
  # major compound groups
  mutate(
    group = `Peak Name` %>% str_remove("\\d+-OH") %>% parse_number() %>% abs() %>% {paste0("C", .)} %>% factor(),
    saturation = case_when(
      str_detect(`Peak Name`, ":0") ~ "n0",
      str_detect(`Peak Name`, ":1") ~ "n1",
      str_detect(`Peak Name`, "cis") ~ "n1 cis",
      str_detect(`Peak Name`, "trans") ~ "n2 trans",
      str_detect(`Peak Name`, ":2") ~ "n2",
      str_detect(`Peak Name`, ":3") ~ "n3",
      str_detect(`Peak Name`, ":4") ~ "n4",
      str_detect(`Peak Name`, ":5") ~ "n5",
      str_detect(`Peak Name`, ":6") ~ "n6",
      str_detect(`Peak Name`, ":7") ~ "n7",
    ),
    is_standard = case_when(
      str_detect(`Peak Name`, "STD") ~ TRUE,
      str_detect(`Peak Name`, "STD", negate = TRUE) ~ FALSE,
    )
  )
```

```{r, fig.width=8, fig.height=13, warning=FALSE}
# Overview Plot
p_fames_db1 <- results %>%
  ggplot() +
  aes(x = `Retention Time`, y = `Relative Height`, fill = saturation, key = `Peak Name`, label = `Peak Name`) +
  geom_bar(stat = 'identity', width = 0.1) +
  facet_wrap(~sample, ncol = 1, scales = "free_y") +
  ggtitle("Soil FAMEs on DB1 Overview") +
  geom_label_repel(aes(vjust =-7, box.padding = 0.5),
                   show.legend = FALSE,
                   size = 2,
                   segment.alpha = 0.5) +
  theme_bw() +
  theme(legend.position = "bottom")
p_fames_db1
```