### 2021-03-08 gc_column_comparison_2.Rmd
image

---
title: "Column Comparison"
author: "Tristan Caro"
date: "2/16/2021"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

# Relevant Equations
## Plates (EP)

$$
N = 5.54\cdot(\frac{R_t}{W})^2
$$
Where N is the plate count (theoretical plates), Rt is retention time, W is peak width at 50% peak height (FWHM).

## Asymmetry (EP)
$$
A_s = \frac{w_{0.05}}{2d}
$$
Where $A_s$ is symmetry factor, $w_{0.05}$ is peak width at one-twentieth of peak height, d is distance between perpendicular dropped from the peak max and the leading edge of the peak at one-twentieth peak height.

## Resolution
Chromeleon uses the European Pharmacopeia (EP) resolution equation.

$$
R = \frac{1.18(Rt_2 - Rt_1)}{(W_2 + W_1)}
$$

Where R is resolution, $Rt$ is retention time, and W is the peak width at 50% peak height (FWHM).

# Libraries

```{r}
# Libraries
library(tidyverse)
library(plotly)
library(ggrepel)
library(ggsci)

# Source
source(file.path("libs", "visualization.R"))
source(file.path("libs", "chromleon.R"))
source(file.path("libs", "chromeleon_peaks.R"))
```

# Load Data

```{r}
# Define path and sample names
root <- "column_testing/run_data"
path_names <- list.files(root)
path_names <- path_names %>% 
  str_subset("blank", negate = TRUE) %>% 
  str_subset("Blank", negate = TRUE) %>% 
  str_subset("GCMS-STD", negate = TRUE)
sample_names <- path_names %>% 
  str_replace("TSQ.....", "") %>% 
  str_replace(".xls", "")

# Create a tibble of integration data
# read_chromeleon_export is a sourced function. See above.
all_data <-
  tibble(
    sample = sample_names,
    file = path_names,
    data = map(file.path(root, file), read_chromeleon_export),
    injection_details = map(data, "injection_details"),
    integration_results = map(data, "integration_results")
  ) %>% select(-data)

# Create a tibble of peak analysis data
peak_data <-
  tibble(
    sample = sample_names,
    file = path_names,
    data = map(file.path(root, file), read_chromeleon_export_peaks),
    injection_details = map(data, "injection_details"),
    integration_results = map(data, "integration_results")
  ) %>% select(-data)
```

```{r, warning = FALSE}
# Data cleanup

results <- all_data %>%
  select(-injection_details) %>% 
  unnest(integration_results) %>% 
  filter(nchar(`Peak Name`) > 0) %>% 
  # major compound groups
  mutate(
    group = `Peak Name` %>% str_remove("\\d+-OH") %>% parse_number() %>% abs() %>% {paste0("C", .)} %>% factor(),
    saturation = case_when(
      str_detect(`Peak Name`, ":0") ~ "n0",
      str_detect(`Peak Name`, "PAIBE") ~ "n0",
      str_detect(`Peak Name`, ":1") ~ "n1",
      str_detect(`Peak Name`, ":2") ~ "n2",
      str_detect(`Peak Name`, ":3") ~ "n3",
      str_detect(`Peak Name`, ":4") ~ "n4",
      str_detect(`Peak Name`, ":5") ~ "n5",
      str_detect(`Peak Name`, ":6") ~ "n6",
      str_detect(`Peak Name`, ":7") ~ "n7",
    ),
    is_standard = case_when(
      str_detect(`Peak Name`, "STD") ~ TRUE,
      str_detect(`Peak Name`, "STD", negate = TRUE) ~ FALSE,
    ),
    column_type = case_when(
      str_detect(`sample`, "DB1") ~ "DB1",
      str_detect(`sample`, "DB5") ~ "DB5",
      str_detect(`sample`, "ZBFAMES") ~ "ZBFAMES",
    ),
    standard_ID = case_when(
      str_detect(`sample`, "BAME") ~ "BAME",
      str_detect(`sample`, "FAMES37") ~ "37FAME",
    )
  )

# Turn Column type into a factor
results$column_type <- factor(results$column_type, levels = c("DB1", "DB5", "ZBFAMES"))

```

```{r, fig.width=10, fig.height=10, warning=FALSE}
# Overview Plot
p_fames_db1 <- results %>%
  ggplot() +
    aes(x = `Retention Time`, y = `Relative Height`, fill = saturation, key = `Peak Name`, label = `Peak Name`) +
    geom_bar(stat = 'identity', width = 0.1) +
    facet_grid(rows = vars(column_type), cols = vars(standard_ID), scales = "free_y") +
    ggtitle("FAME Standards Overview") +
    #geom_label_repel(aes(vjust =-7, box.padding = 0.5),
                     #show.legend = FALSE,
                     #size = 2,
                     #segment.alpha = 0.5) +
    theme_bw() +
    scale_fill_npg() +
    theme(legend.position = "bottom")
p_fames_db1
```