### 2021-07-20 soil_report.Rmd
Soil analysis Rmd

---
title: "Soil Quality"
author: "Tristan Caro"
date: "7/20/2021"
output: html_document
---

# Soil Quality Report
Soils were submitted to CSU Soil Testing Lab on 6/10/2021. Data was reported on 6/30/2021.
The lab tests for pH, salts, excess lime, texture, organic matter, as well as N, P, K, Zn, Fe, Mn, Cu, and S.

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(ggsci)
```

```{r, message=FALSE}
soil_data <- read_csv("soil_report_06302021.csv")

soil_qualities <- soil_data %>% 
  select(c(`Sample ID`, `Excess Lime`, `Texture Estimate`))

soil_values <- soil_data %>% 
  select(!c(`Excess Lime`, `Texture Estimate`))

soil_values_long <- soil_values %>% 
  select(!`Sample Depth (cm)`) %>% 
  pivot_longer(!`Sample ID`, names_to = "Measurement", values_to = "Value")

```

```{r}
p_soil_values_long <- soil_values_long %>% 
  ggplot(aes(x = `Sample ID`, y = `Value`, fill = `Sample ID`)) +
  geom_col() +
  facet_wrap(~`Measurement`, scales = "free_y") +
  scale_fill_aaas() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("soil_report_plot.png", plot = p_soil_values_long, height = 10, width = 15)
```