### 2021-08-11 gc_ramp_comparison.Rmd
tidy data and overview plot

---
title: "R Notebook"
output: html_notebook
---

```{r, include=FALSE}
# Libraries
library(tidyverse) # What are we without the tidyverse? c:
library(plotly)
library(ggrepel)
library(ggsci)
library(ggsignif)

# Source
source(file.path("libs", "visualization.R"))
source(file.path("libs", "chromleon.R"))
source(file.path("libs", "chromeleon_peaks.R"))
```

# Relevant Equations

## Plates (EP)
Chromeleon uses the European Pharmacopeia (EP) theoretical plate calculation.

$$
N = 5.54\cdot(\frac{R_t}{W})^2
$$
Where N is the plate count (theoretical plates), Rt is retention time, W is peak width at 50% peak height (FWHM).

## Asymmetry (EP)
$$
A_s = \frac{w_{0.05}}{2d}
$$
Where $A_s$ is symmetry factor, $w_{0.05}$ is peak width at one-twentieth of peak height, d is distance between perpendicular dropped from the peak max and the leading edge of the peak at one-twentieth peak height.

## Resolution
Chromeleon uses the European Pharmacopeia (EP) resolution equation.

$$
R = \frac{1.18(Rt_2 - Rt_1)}{(W_2 + W_1)}
$$

Where R is resolution, $Rt$ is retention time, and W is the peak width at 50% peak height (FWHM).

# Load Data
```{r}
root <- "data/2021-08-10_TAC_BAME_Ramp_Tests/"
# Generate a list of files in the directory
file_list = list.files(path = root)
# Subset this list for the samples we want
file_list <- file_list %>% 
  str_subset("blank", negate = TRUE) %>% # remove blanks!
  str_subset("Blank", negate = TRUE) %>% # remove more blanks!
  str_subset("GCMS-STD", negate = TRUE) %>% # remove GCMS Standards!
  str_subset("BLK", negate = TRUE) # remove even more blanks! ^_^
# Make a list of sample names without all the mumbo jumbo in the exact filename
sample_names <- file_list %>% 
  str_replace("FID.....", "") %>% # remove FID header
  str_replace("_100ul-nhex.xls", "") %>% # remove volume and type of solvent
  str_replace("_100ul.xls", "") %>% 
  str_replace(".xls", "") %>% 
  str_replace("TCBAME6_", "")

# Make a big tibble containing all of our data
# With Injection Details and Integration Results nested inside
all_data <-
  tibble(
    sample = sample_names,
    file = file_list,
    data = map(file.path(root, file), read_chromeleon_export),
    injection_details = map(data, "injection_details"),
    integration_results = map(data, "integration_results")
  ) %>% select(-data)

# Create a tibble of peak analysis data
peak_data <-
  tibble(
    sample = sample_names,
    file = file_list,
    data = map(file.path(root, file), read_chromeleon_export_peaks),
    injection_details = map(data, "injection_details"),
    integration_results = map(data, "integration_results")
  ) %>% select(-data)

# Make a tibble containing all of the integration data
peak_analysis <- peak_data %>%
  select(-injection_details) %>% 
  unnest(integration_results) %>% 
  filter(nchar(`Peak Name`) > 0)

results <- all_data %>%
  select(-injection_details) %>% 
  unnest(integration_results) %>% 
  filter(nchar(`Peak Name`) > 0) %>% 
  filter(sample != "TCBAME-2") %>% # Don't need this sample
  # Add ramp information
  mutate(ramp = str_extract(sample, "\\d"))
```

# Overview Plot
```{r, fig.width=8, fig.height=4, warning=FALSE}
# Overview Plot
p_bame_overview <- results %>%
  ggplot() +
    aes(x = `Retention Time`, y = `Relative Height`, fill = ramp) +
    geom_col(width = 0.1) +
    facet_wrap(.~ramp, ncol = 1) +
    ggtitle("FAME Standards Overview") +
    theme_bw() +
    scale_fill_npg() +
    theme(legend.position = "bottom")
p_bame_overview
```