### 2021-09-01 irms_ramp_testing/irms_ramp_testing.Rmd
irms_ramp_testing

---
title: "IRMS Ramp Tests"
author: "Tristan Caro"
date: "8/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(ggdist)
```

# Read in the data
```{r}
file_list <- list.files("data/", pattern='*.xls')
ramp_data <- do.call(rbind, lapply(paste0("data/", file_list), read_excel))

res_calc <- function(rt2, rt1, w2, w1) {
  res = (rt2 - rt1) / (1/2)*(w1 + w2)
  return(res)
}

# Clean up some <chr>
ramp_data$`Area 2 (Vs)` <- as.numeric(ramp_data$`Area 2 (Vs)`)
ramp_data$`Ampl  2 (mV)` <- as.numeric(ramp_data$`Ampl  2 (mV)`)

# Make Peak Numbers Numeric and filter out gas calibration peaks
ramp_data <- ramp_data %>%
  mutate(`Peak Nr` = parse_number(`Peak Nr`)) %>% 
  group_by(Filename) %>% 
  # Remove first 10 peaks and last 5 that correspond to gas calibration
  filter(`Peak Nr` > 10, `Peak Nr` < max(`Peak Nr`) - 5)

```

Calculate resolution using the following:

$$
R = \frac{t_{R2} - t_{R1}}{\frac{1}{2}(W_1 + W_2)}
$$

# BAMES
```{r}
bames <- ramp_data %>% 
  filter(str_detect(Filename, "BAMES")) %>% 
  mutate(ramp_cpm = case_when(
    # Not the best way to assign ramp temp .. brain dead.
    str_detect(Filename, "5_0Cpm") ~ "5",
    str_detect(Filename, "2_0Cpm") ~ "2.0",
    str_detect(Filename, "1_5Cprm") ~ "1.5",
    str_detect(Filename, "1_0Cpm") ~ "1.0")
  ) %>% 
  group_by(Filename) %>% 
  mutate(rt2 = as.numeric(`Rt (s)`), # later peak
         rt1 = as.numeric(lag(`Rt (s)`, n = 1)), # earlier peak
         w2 = as.numeric(`Width (s)`), # later peak
         w1 = as.numeric(lag(`Width (s)`, n = 1)), # earlier peak
         resolution = res_calc(rt2 = rt2, rt1 = rt1, w2 = w2, w1 = w1))
```

```{r}
bames %>% ggplot(aes(x = `Filename`, y = resolution, color = ramp_cpm)) +
  stat_pointinterval(color = "black") +
  geom_jitter(width = 0.2, alpha = 0.3) +
  theme_classic() +
  ggtitle("BAMES Resolution vs. Ramp") +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank())
```

```{r}
bames %>%
  ggplot(aes(x = `Area 2 (Vs)`, y = resolution, color = ramp_cpm)) +
  geom_point() +
  geom_smooth(se = FALSE,
              method = "lm") +
  facet_wrap(~ramp_cpm) +
  theme_classic()
```

```{r}
bames %>%
  ggplot(aes(x = `Ampl  2 (mV)`, y = resolution, color = ramp_cpm)) +
  geom_point() +
  geom_smooth(se = FALSE,
              method = "lm") +
  facet_wrap(~ramp_cpm) +
  theme_classic()
```

# Soil FAME sample t0f0x

```{r}
sfames <- ramp_data %>% 
  filter(str_detect(Filename, "t0f0x")) %>% 
  mutate(ramp_cpm = case_when(
    # Not the best way to assign ramp temp .. brain dead.
    str_detect(Filename, "5_0Cpm") ~ "5",
    str_detect(Filename, "2_0Cpm") ~ "2.0",
    str_detect(Filename, "1_5Cpm") ~ "1.5",
    str_detect(Filename, "1_0Cpm") ~ "1.0")
  ) %>% 
  group_by(Filename) %>% 
  mutate(rt2 = as.numeric(`Rt (s)`), # later peak
         rt1 = as.numeric(lag(`Rt (s)`, n = 1)), # earlier peak
         w2 = as.numeric(`Width (s)`), # later peak
         w1 = as.numeric(lag(`Width (s)`, n = 1)), # earlier peak
         resolution = res_calc(rt2 = rt2, rt1 = rt1, w2 = w2, w1 = w1))
```

```{r}
sfames %>% 
  ggplot(aes(x = `Filename`, y = resolution, color = ramp_cpm)) +
  stat_pointinterval(color = "black") +
  geom_jitter(width = 0.2, alpha = 0.3) +
  coord_cartesian(ylim = c(0, 100000)) +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank())
```

```{r}
sfames %>%
  ggplot(aes(x = `Area 2 (Vs)`, y = resolution, color = ramp_cpm)) +
  geom_point() +
  theme_classic()
```

```{r}
sfames %>%
  ggplot(aes(x = `Ampl  2 (mV)`, y = resolution, color = ramp_cpm)) +
  geom_point() +
  theme_classic()
```

```{r}
sfames %>%
  ggplot(aes(x = `Ampl  2 (mV)`, y = resolution, color = ramp_cpm)) +
  geom_point() +
  facet_wrap(~ramp_cpm, scales = "free") +
  theme_classic()
```