### 2022-01-07 soil_report.Rmd
cleaned up soil report

---
title: "Soil Quality"
author: "Tristan Caro"
date: "7/20/2021"
output: html_document
---

# Soil Quality Report
Soils were submitted to CSU Soil Testing Lab on 6/10/2021. Data was reported on 6/30/2021.
The lab tests for pH, salts, excess lime, texture, organic matter, as well as N, P, K, Zn, Fe, Mn, Cu, and S.

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(ggsci)
```

```{r, message=FALSE}
soil_data <- read_csv("data/soil_data/soil_report_06302021.csv")

soil_data <- soil_data %>% 
  mutate(
    soil_type = case_when(
      str_detect(`Sample ID`, "grassland") ~ "Grassland",
      str_detect(`Sample ID`, "tundra") ~ "Tundra",
      str_detect(`Sample ID`, "confier") ~ "Conifer Forest",
      str_detect(`Sample ID`, "meadow_0cm") ~ "Meadow 0cm",
      str_detect(`Sample ID`, "meadow_30cm") ~ "Meadow 30cm"
    )
  )

soil_qualities <- soil_data %>% 
  select(c(`Sample ID`, `Excess Lime`, `Texture Estimate`))

soil_values <- soil_data %>% 
  select(!c(`Excess Lime`, `Texture Estimate`, `Sample ID`))

soil_values_long <- soil_values %>% 
  select(!`Sample Depth (cm)`) %>% 
  pivot_longer(!soil_type, names_to = "Measurement", values_to = "Value")

```

```{r}
p_soil_values_long <- soil_values_long %>% 
  filter(soil_type %in% c("Tundra", "Grassland", "Conifer Forest")) %>% 
  ggplot(aes(x = soil_type, y = `Value`, fill = soil_type)) +
  geom_col() +
  facet_wrap(~`Measurement`, scales = "free_y") +
  scale_fill_futurama() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p_soil_values_long

#ggsave("soil_report_plot.png", plot = p_soil_values_long, height = 10, width = 15)
```