### 2022-03-23 00_LH-SIP_Data_Reduction.Rmd
separated data reduction from qSIP script

---
title: "00_LH-SIP_Data_Reduction"
author: "Tristan Caro"
date: "`r Sys.Date()`"
output:
    html_document:
    code_folding: show
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---

# Setup

```{r setup, include = FALSE}
# global knitting options for code rendering
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf"),
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.keep = "all",
  fig.path = file.path("fig_output", paste0(gsub("\\.[Rr]md", "", knitr::current_input()), "_"))
)
```

## Load packages

```{r, message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)    # CRAN v1.3.1
library(forcats)      # CRAN v0.5.1 
library(isoreader)    # [::NA/NA] v1.3.0 # CRAN v1.3.0
library(isoprocessor) # [github::isoverse/isoprocessor] v0.6.7
library(isotopia)     # [github::isoverse/isotopia] v0.5.8 
library(readxl)       # CRAN v1.3.1 
library(ggsci)        # CRAN v2.9 
library(ggborderline) # [github::wurli/ggborderline] v0.1.0
library(ggrepel)      # CRAN v0.9.1 
library(latex2exp)    # CRAN v0.5.0
library(ggdist)       # CRAN v3.0.0 
library(ggsignif)     # CRAN v0.6.2
library(ggridges)     # CRAN v0.5.3
library(cowplot)      # CRAN v1.1.1
```

## Load sourced functions
These functions are found in the `libs/` directory.
```{r}
# Sourced Functions
source(file.path("libs", "visualization.R"))              # Visualization scripts
source(file.path("libs", "chromeleon.R"))                 # Chromeleon Reader
source(file.path("libs", "error_prop.R"))                 # Error Propagation
source(file.path("libs", "calculate_turnover.R"))         # Turnover Calculator
`%nin%` = Negate(`%in%`)                                  # "Not In" function
var_to_str = function(v) {return(deparse(substitute(v)))} # Convert variable name to string
```

# Load data

## Load F_L 2F Data

```{r}
F_L_2F <- read_csv("data/Label_2H_Data/dD_cleaned.csv") %>% 
  group_by(soil) %>% 
  summarize(F_label_ppm_mn = mean(F_label_ppm),
            F_label_ppm_se = sd(F_label_ppm)) %>% 
  mutate(F_label_mn = F_label_ppm_mn / 10000,
         F_label_se = F_label_ppm_se / 10000) %>% 
  mutate(terrain = case_when(
    soil == "Conifer" ~ "Gordon Gulch Conifer Forest",
    soil == "Tundra" ~ "Niwot Ridge Tundra",
    soil == "Grassland" ~ "Marshall Mesa Grassland"
  ))
```

## Load Compound nC and nH Data sheet

```{r}
compound_nC_nH <- read_xlsx("data/compound_nC_nH.xlsx")
```

## Load Phthalic Acid IRMS Data

```{r}
# Load PAME Iso Files
PAME_data_path <- file.path("data", "IRMS", "PAME_data")
ref_ratio <- get_standard("2H") %>% as.numeric()

# read files
PAME_iso_files_raw <- 
  # path to data files
  PAME_data_path %>% 
  # read data files in parallel for fast read
  iso_read_continuous_flow() %>%
  # filter out files with read errors (e.g. from aborted analysis)
  iso_filter_files_with_problems()

# Load Peak Map

PAME_peak_map <- 
  # initial peak map:
  # readxl::read_excel(file.path("data","IRMS", "peak_map_manual_general.xlsx"))
  # corrected peak map
  readxl::read_excel(file.path("data", "IRMS", "PAME_peak_map.xlsx"))
PAME_peak_map

# process file information
PAME_iso_files <- PAME_iso_files_raw %>% 
  # rename key file info columns
  iso_rename_file_info(analysis = Analysis, id1 = `Identifier 1`, id2 = `Identifier 2`) %>% 
  # parse text info into numbers
  iso_parse_file_info(number = analysis) %>% 
  # process other file information that is specific to the naming conventions
  # of this particular sequence
  iso_mutate_file_info(
    # what is the type of each analysis?
    type = case_when(
      str_detect(id1, "[Zz]ero")      ~ "on_off",
      str_detect(id1, "H3")           ~ "H3_factor",
      str_detect(id1, "F8")           ~ "F8_std",
      str_detect(id1, "F9")           ~ "F9_std",
      TRUE                            ~ "sample"
    ),
    # what was the concentration? (assuming Preparation = concentration or volume)
    concentration = 
      ifelse(type == "std",  
             str_extract(Preparation, "[0-9.]+ ?ng( per |/)uL") %>% 
               parse_number() %>% iso_double_with_units("ng/uL"),
             NA),
    # what folder are the data files in? (assuming folder = sequence)
    folder = basename(dirname(file_path))
  ) %>% 
  # focus only on the relevant file info, discarding the rest
  iso_select_file_info(
    folder, analysis, file_datetime, id1, type, concentration
  )

# Peak Mapping
# set peak table from vendor data table with default isodat template
PAME_iso_files <- iso_set_peak_table_from_isodat_vendor_data_table(PAME_iso_files) %>% 
  # convert units from mV to V for amplitudes and area
  iso_convert_peak_table_units(V = mV, Vs = mVs)

# Create Peak Table and filter out non-PAME samples
PAME_peak_table <- PAME_iso_files %>%
  iso_filter_files(str_detect(id1, "PAME")) %>% 
  # Choose only runs that worked
  iso_filter_files(str_detect(file_id, "success")) %>% 
  iso_set_peak_table_from_isodat_vendor_data_table() %>%
  iso_get_peak_table() %>% 
  mutate(Analysis = substr(file_id, 1, 7),
         sample_id = substr(file_id, 10,14))

PAME_peak_mapped <- PAME_peak_table %>%
  iso_map_peaks(PAME_peak_map, map_id = Analysis) %>%
  filter(!is.na(compound))

# Summarize the mean and standard error
PAME_summarized <- PAME_peak_mapped %>%
  summarize(at2H_mn = mean(at2H),
            at2H_se = sd(at2H)) %>% 
  mutate(at2H_mn_ppm = at2H_mn * 10000,
         at2H_se_ppm = at2H_se * 10000)
```

## Load IRMS Data

### Load Metadata and Problematic Run Data

```{r}
mtda <- read_csv(file.path("data", "sample_metadata_2021.csv"))
sample_names <- mtda$id1

problematic_runs <- read_xlsx(file.path("data", "IRMS", "problematic_runs.xlsx"))
problematic_analyses <- problematic_runs$Analysis
problematic_analyses_num <- parse_number(problematic_analyses)
```

### Load Peak Map

```{r}
# this information is often maintained in a csv or Excel file
peak_map <- 
  # initial peak map:
  # readxl::read_excel(file.path("data","IRMS", "peak_map_manual_general.xlsx"))
  # corrected peak map
  readxl::read_excel(file.path("data", "IRMS", "peak_maps_resolved.xlsx"))
```

### Read raw IRMS data files

```{r, warning=FALSE}
# Set file path(s) to data files, folders or rds collections 
# can be multiple folders or mix of folders and files
# Isoverse will iteratively search subfolders. Huzzah! ^_^
data_path <- file.path("data", "IRMS", "raw_data")
ref_ratio <- get_standard("2H") %>% as.numeric()

# read files
iso_files_raw <- 
  # path to data files
  data_path %>% 
  # read data files in parallel for fast read
  iso_read_continuous_flow() %>%
  # filter out files with read errors (e.g. from aborted analysis)
  iso_filter_files_with_problems()
```

### Process file info & peak table

```{r}
# process IRMS file information
iso_files <- iso_files_raw %>% 
  # rename key file info columns
  iso_rename_file_info(analysis = Analysis, id1 = `Identifier 1`, id2 = `Identifier 2`) %>% 
  # parse text info into numbers
  iso_parse_file_info(number = analysis) %>% 
  # process other file information that is specific to the naming conventions
  # of this particular sequence
  iso_mutate_file_info(
    # what is the type of each analysis?
    type = case_when(
      str_detect(id1, "[Zz]ero")      ~ "on_off",
      str_detect(id1, "H3")           ~ "H3_factor",
      str_detect(id1, "F8")           ~ "F8_std",
      str_detect(id1, "F9")           ~ "F9_std",
      TRUE                            ~ "sample"
    ),
    # what was the concentration? (assuming Preparation = concentration or volume)
    concentration = 
      ifelse(type == "std",  
             str_extract(Preparation, "[0-9.]+ ?ng( per |/)uL") %>% 
               parse_number() %>% iso_double_with_units("ng/uL"),
             NA),
    # what folder are the data files in? (assuming folder = sequence)
    folder = basename(dirname(file_path))
  ) %>% 
  # focus only on the relevant file info, discarding the rest
  iso_select_file_info(
    folder, analysis, file_datetime, id1, type, concentration
  ) %>% 
  # add in additional sample metadata (could be any info)
  # note: this would typically be stored in / read from a csv or excel file
  iso_add_file_info(
    read_csv(file.path("data", "sample_metadata_2021.csv")),
    join_by = "id1"
  )

# set peak table from vendor data table with default isodat template
iso_files <- iso_set_peak_table_from_isodat_vendor_data_table(iso_files) %>% 
  # convert units from mV to V for amplitudes and area
  iso_convert_peak_table_units(V = mV, Vs = mVs)

# focus on sample files
sample_files <- iso_filter_files(iso_files, type == "sample")

# Include only sample files
sample_files <- iso_filter_files(sample_files, id1 %in% sample_names)

# EXCLUDE CRC SAMPLES FROM THIS ANALYSIS
sample_files <- iso_filter_files(sample_files, str_detect(id1, "cri", negate = TRUE))

sample_files <- iso_filter_files(sample_files, str_detect(id1, "cr0", negate = TRUE))

# EXCLUDE RUN WE'VE DEEMED PROBLEMATIC
sample_files <- iso_filter_files(sample_files, analysis %nin% problematic_analyses_num)

```

### Show file information

```{r}
# display file information
iso_files %>% 
  iso_get_file_info() %>% select(-file_id, -folder) %>% 
  iso_make_units_explicit() %>% knitr::kable()
```

### Example chromatogram

```{r "example_chromatograms", fig.width=8, fig.height=8}
# plot an example chromatogram

# sample_files[10] %>% # choosing arbitrary file to plot
# iso_plot_continuous_flow_data(
#   # select data and aesthetics
#   data = c(2), color = id1, panel = id1,
#   # zoom in on time interval
#   time_interval = c(750, 4000),
#   # peak labels for all peaks > 2V
#   peak_bounds = TRUE,
#   peak_marker = FALSE,
#   peak_label = iso_format(rt),
#   #peak_label_size = 3,
#   peak_label_filter = analysis == 5685 & amp2 > 1
# ) + scale_color_npg() + 
#   theme_classic() +
#   theme(strip.background = element_blank())
```

## Load GC-FID Data

```{r}
root <- file.path("data", "gc_fid_data_2021")
file_list = list.files(path = root)
file_list <- file_list %>% 
  str_subset("blank", negate = TRUE) %>% 
  str_subset("Blank", negate = TRUE) %>% 
  str_subset("GCMS-STD", negate = TRUE) %>% 
  str_subset("BLK", negate = TRUE) # remove procedural blanks
sample_names <- file_list %>% 
  str_replace("FID.....", "") %>% 
  str_replace("_100ul-nhex.xls", "") %>% 
  str_replace("_100ul.xls", "")

FID_all_data <-
  tibble(
    sample = sample_names,
    file = file_list,
    data = map(file.path(root, file), read_chromeleon_export),
    injection_details = map(data, "injection_details"),
    integration_results = map(data, "integration_results")
  ) %>% select(-data)
```

### Unnest FID integration details and clean

```{r}
FID_results <- FID_all_data %>% 
  select(-injection_details) %>% 
  unnest(integration_results) %>% 
  filter(nchar(`Peak Name`) > 0) %>% 
  # Specify chain length
  mutate(
    chain = `Peak Name` %>% 
      str_remove("\\d+-OH") %>% 
      parse_number() %>% 
      abs() %>% 
      {paste0("C", .)} %>% 
      factor()
      ) %>% 
  # Rename these columns so we can join the IRMS and FID data
  # Important to specify what data comes from FID!
  rename(
    "sample_id" = sample,
    "compound" = `Peak Name`,
    "rt_FID" = `Retention Time`,
    "area_FID" = `Area`,
    "height_FID" = `Height`,
    "rel_area_FID" = `Relative Area`,
    "rel_height_FID" = `Relative Height`,
    "amount_FID" = `Amount`
         ) %>% 
  # Make FID notation match IRMS notation
  mutate(
    compound = 
      case_when(
        str_detect(compound, "18:2 trans 9, 12") ~ "18:2 9, 12",
        str_detect(compound, "18:2 trans 9,12") ~ "18:2 9, 12",
        str_detect(compound, "18:2 9,12") ~ "18:2 9, 12",
        str_detect(compound, "16:1 cis-9") ~ "16:1 cis 9",
        str_detect(compound, "16:1 trans-9") ~ "16:1 trans 9",
        str_detect(compound, "18:1 cis-9") ~ "18:1 cis 9",
        str_detect(compound, "18:1 trans-9") ~ "18:1 trans 9",
        str_detect(compound, "21:0") ~ "21:0 (STD)",
        str_detect(compound, "23:0") ~ "23:0 (STD)",
        TRUE ~ compound
      )
  )
```

# Compound abundance

```{r compound_abundance}
# Compute the instrument response for each run:
# Response = µg per unit area using 10µg of PAIBE as quantification STD

FID_results_PAIBE_ug <- FID_results %>% 
  select(
    sample_id, 
    file,
    compound,
    area_FID, 
    rel_area_FID) %>% 
  filter(compound == "PAIBE") %>% 
  mutate(
    PAIBE_area = area_FID,
    PAIBE_ug_inj = 0.1, # PAIBE injected = (10µg/100µL*1µL) = 0.1µg
    # Find instrument response as PAIBE mass per FID area
    ug_per_area = PAIBE_ug_inj / PAIBE_area,
    # Check that the inferred mass is the same as known mass
    PAIBE_ug_inj_inferred = ug_per_area * PAIBE_area
  ) %>%
  select(file, PAIBE_area, PAIBE_ug_inj, ug_per_area)

```

## Calculate mass of analyte injected into GC-FID

```{r response}
# Put the response info back into the abundance data frame:
FID_results_abundance <- 
  left_join(
    FID_results, 
    FID_results_PAIBE_ug, 
    by = "file"
    ) %>%
  # Calculate the quantity of analyte injected based off of area and 
  # instrument response
  mutate(
    compound_ug_inj = area_FID * ug_per_area,
    inj_vol_ul = 1, # GC-FID injection volume µL
    susp_vol_ul = 100, # suspension volume µL
    # The mass of analyte in vial is the concentration of analyte (mass 
    # seen per unit volume injected) times the suspension volume.
    compound_ug_vial = (compound_ug_inj * inj_vol_ul) * susp_vol_ul
  )

```

## Calculate fractional derivatization efficiency (21-PC)
```{r 21_PC}
FID_results_abundance_21 <- FID_results_abundance %>% 
  filter(compound == "21:0 (STD)") %>% 
  # Derivatization efficiency is the amount of standard observed divided by 
  # the amount of standard added (21:0 PC = 10µg)
  mutate(
    STD21_ug = 10,
    deriv_efficiency = compound_ug_vial / STD21_ug
  ) %>% 
  select(file, STD21_ug, deriv_efficiency)

# Add deriv efficiency back into results data frame
FID_results_abundance <- 
  left_join(
    FID_results_abundance, 
    FID_results_abundance_21, 
    by = "file"
  ) %>% 
  # Calculate mass of fatty acids in extraction mixture as
  # compounds present as FAMEs multiplied by derivatization efficiency
  mutate(compound_ug_ext = compound_ug_vial / deriv_efficiency)
```

## Calculate extraction efficiency (23-PC)

```{r 23_PC}
FID_results_abundance_23 <- FID_results_abundance %>% 
  filter(compound == "23:0 (STD)") %>% 
  # Lipid Recovery Efficiency is the amount of initial standard 
  # observed divided by the amount of standard added (23:0 PC = 100ug)
  mutate(
    STD23_ug = 100,
    extr_efficiency = compound_ug_vial / STD23_ug
  ) %>% 
  select(file, STD23_ug, extr_efficiency)

# Add extr efficiency back into results data frame
FID_results_abundance <- 
  left_join(
    FID_results_abundance,
    FID_results_abundance_23,
    by = "file"
  ) %>% 
  mutate(
    # Calculate amount of analyte in soil sample
    compound_ug_soil_sample = compound_ug_ext / extr_efficiency,
    # Calculate amount of analyte per gram of soil (3g was used for each extraction)
    soil_mass_g = 10, # 10 g of each soil were used for this study
    compound_ug_per_g_soil = compound_ug_soil_sample / soil_mass_g
  )
```

## Remove standards from soil abundances

```{r}
# Remove standards from soil abundances

FID_results_abundance_no_stds <- FID_results_abundance %>% 
  filter(
    compound %nin% c("21:0 (STD)", "23:0 (STD)", "PAIBE"),
    str_detect(sample_id, "FAME", negate = TRUE),
    str_detect(sample_id, "BAME", negate = TRUE)
  )
```

# IRMS Peak Mapping

```{r}
sample_peak_table <- sample_files %>%
  iso_set_peak_table_from_isodat_vendor_data_table() %>%
  iso_get_peak_table() %>% 
  mutate(Analysis = substr(file_id, 1, 7),
         sample_id = substr(file_id, 10,14))
```

## Generate Isoverse Peak Map
```{r}
# Generate an isoverse peak map
peaks_mapped <- sample_peak_table %>%
  iso_map_peaks(peak_map, map_id = Analysis) %>%
  filter(!is.na(compound))
```

## Remove readychecks, blanks, etc. from analysis
```{r}
# Filter based on what is in the sample_names metadata
peaks_mapped <- peaks_mapped %>% 
  filter(sample_id %in% sample_names)

# Filter out runs that are on our problematic_runs list
peaks_mapped <- peaks_mapped %>% 
  filter(Analysis %nin% problematic_runs)
```

## Check problematic peak assignments
```{r}
problem_peaks <- iso_get_problematic_peak_mappings(peaks_mapped)
```

## Export summary of unresolved peaks
```{r}
# Spit out peaks that need to be resolved
peaks_mapped %>%
  ungroup() %>%
  select(Analysis, compound, rt) %>%
  mutate(rt = round(rt, digits=0)) %>%
  pivot_wider(
    names_from = Analysis, 
    names_prefix = "rt:",
    values_from = rt, 
    id_cols = compound, 
    values_fn = function(x) paste(x, collapse = "; ")
  ) %>%
  openxlsx::write.xlsx(file.path("data", "IRMS", "peak_maps_resolve.xlsx"))
```

## Assign incubation params
```{r}
# Add Some Parameters
peaks_mapped <- peaks_mapped %>% 
  group_by(sample_id) %>% 
  mutate(inc_time_d = as.numeric(substr(sample_id, 4,4)),
         inc_time_d_str = paste(inc_time_d, "Days"),
         f_label = 0.005,
         t_series_id = paste0(substr(sample_id, 1,1), 
                              substr(sample_id, 5,5)),
         soil_id = substr(t_series_id, 1,1),
         replicate_id = substr(t_series_id, 2,2),
         terrain = case_when(
           substr(sample_id, 1, 1) == "t" ~ "Niwot Ridge Tundra",
           substr(sample_id, 1 ,1) == "c" ~ "Gordon Gulch Conifer Forest",
           substr(sample_id, 1 ,1) == "m" ~ "Gordon Gulch Meadow",
           substr(sample_id, 1 ,1) == "g" ~ "Marshall Mesa Grassland"))
# 
# inc_time_zero <- peaks_mapped %>% 
#   filter(inc_time_d == 0) %>% 
#   select(sample_id, compound, at2H, t_series_id) %>% 
#   mutate(f_start = at2H) %>% 
#   select(-at2H)
# 
#

```

## Extract zero timepoint values
```{r}
zero_peaks <- peaks_mapped %>%
  filter(inc_time_d_str == "0 Days") %>%
  group_by(t_series_id, compound) %>%
  mutate(n_analyical_reps = n()) %>%
  ungroup() %>%
  # exclude ambiguous peaks
  filter(!is_ambiguous)

zero_peaks_averaged <- zero_peaks %>% 
  group_by(soil_id, compound) %>% 
  summarise(at2H_mn_zero = mean(at2H),
            d2H_mn_zero = mean(d2H),
            area2_mn_zero = mean(area2))

zero_peaks_averaged %>% 
  openxlsx::write.xlsx(file.path("data", "IRMS", "zero_peaks_averaged.xlsx"),
                       overwrite = TRUE)

# zero_peaks_averaged <- read_xlsx(file.path("data", "IRMS", "zero_peaks_averaged.xlsx"))

# This is the dataset we will use for generation time calculations
peaks_mapped_with_zeros <- 
  peaks_mapped %>%
  left_join(
    zero_peaks_averaged,
    by = c("soil_id", "compound")
  )
stopifnot(nrow(peaks_mapped_with_zeros) == nrow(peaks_mapped))

# no zero samples
# peaks_mapped_with_zeros %>% filter(is.na(zero_sample_id)) %>% View()

peaks_mapped_with_zeros <- peaks_mapped_with_zeros %>% 
  mutate(
    sample_id = substr(file_id, 10,14)
    )
```

## Extract 3 day timepoint values

```{r}
three_day_peaks <- peaks_mapped %>%
  filter(inc_time_d_str == "3 Days") %>%
  group_by(t_series_id, compound) %>%
  mutate(n_analyical_reps = n()) %>%
  ungroup() %>%
  # exclude ambiguous peaks
  filter(!is_ambiguous)

three_day_peaks_averaged <- three_day_peaks %>% 
  group_by(soil_id, compound) %>% 
  summarise(at2H_mn_three = mean(at2H),
            d2H_mn_three = mean(d2H),
            area2_mn_three = mean(area2))

peaks_mapped_with_threes <- 
  peaks_mapped %>%
  left_join(
    three_day_peaks_averaged,
    by = c("soil_id", "compound")
  )
stopifnot(nrow(peaks_mapped_with_threes) == nrow(peaks_mapped))

peaks_mapped_with_threes <- peaks_mapped_with_threes %>% 
  mutate(
    sample_id = substr(file_id, 10,14)
    )
```

## Join IRMS and GC-FID Data

```{r}
# peaks_mapped_with_zeros_FID <- peaks_mapped_with_zeros %>% 
#   left_join(FID_results, by = c("sample_id", "compound"))

peaks_mapped_with_zeros_FID <- peaks_mapped_with_zeros %>%
  left_join(
    FID_results_abundance_no_stds, 
    by = c("sample_id", "compound")
  )

peaks_mapped_with_threes_FID <- peaks_mapped_with_threes %>% 
    left_join(
    FID_results_abundance_no_stds, 
    by = c("sample_id", "compound")
  )
```

## Correct for PAME

```{r}

# For 0 -3, 0-7 timepoints
peaks_mapped_with_zeros_FID <- peaks_mapped_with_zeros_FID %>% 
  left_join(compound_nC_nH, by = "compound") %>% 
  mutate(`2F_Me` = PAME_summarized %>% pull(at2H_mn),
         `2F_FAME` = at2H,
         `2F_alk` = (`2F_FAME` - `2F_Me`*x_me) / x_alk)

# for 3-7 timepoints
peaks_mapped_with_threes_FID <- peaks_mapped_with_threes_FID %>% 
  left_join(compound_nC_nH, by = "compound") %>% 
  mutate(`2F_Me` = PAME_summarized %>% pull(at2H_mn),
         `2F_FAME` = at2H,
         `2F_alk` = (`2F_FAME` - `2F_Me`*x_me) / x_alk)
```

## Separate analytes and standards, calculate F_0 se

```{r}
# For 0-3,0-7 timepoints
# Analytes
peaks_mapped_with_zeros_analytes <- peaks_mapped_with_zeros_FID %>% 
  # Don't want our standard compounds to be used for generation time calculations! x_x
  filter(str_detect(compound, "STD", negate = TRUE)) %>% 
  # Also don't want our yeast-extract stimulated condition in our analyte pool
  filter(str_detect(sample_id, "t0t", negate = TRUE)) %>% 
  left_join(F_L_2F, by = "terrain")

# Standards
peaks_mapped_with_zeros_standards<- peaks_mapped_with_zeros_FID %>% 
  filter(str_detect(compound, "STD"))

# 23:0 and 21:0 PC standard errors at inc time zero
# Determines our standard error in FAME measurements at F_0
sF_0 <- peaks_mapped_with_zeros_standards %>% 
  # Select only 21 and 23 PC standards at initial time point
  filter(compound %in% c("21:0 (STD)", "23:0 (STD)"),
         inc_time_d == 0) %>%
  # Calculate mean and se of all 21:0 and 23:0 standard at2H values
  ungroup() %>% 
  summarize(F_0_at2H_mn = mean(at2H),
            F_0_at2H_se = sd(at2H))

# Add sF_0 to analyte data

peaks_mapped_with_zeros_analytes <- peaks_mapped_with_zeros_analytes %>% 
  mutate(F_0_at2H_se = 
           sF_0 %>% pull(F_0_at2H_se)
         )

# For 3 -7 timepoints
# Analytes
peaks_mapped_with_threes_analytes <- peaks_mapped_with_threes_FID %>% 
  # Don't want our standard compounds to be used for generation time calculations! x_x
  filter(str_detect(compound, "STD", negate = TRUE)) %>% 
  # Also don't want our yeast-extract stimulated condition in our analyte pool
  filter(str_detect(sample_id, "t0t", negate = TRUE)) %>% 
  left_join(F_L_2F, by = "terrain")

# Standards
peaks_mapped_with_threes_standards<- peaks_mapped_with_threes_FID %>% 
  filter(str_detect(compound, "STD"))

# 23:0 and 21:0 PC standard errors at inc time zero
# Determines our standard error in FAME measurements at F_0
sF_0 <- peaks_mapped_with_threes_standards %>% 
  # Select only 21 and 23 PC standards at initial time point
  filter(compound %in% c("21:0 (STD)", "23:0 (STD)"),
         inc_time_d == 0) %>%
  # Calculate mean and se of all 21:0 and 23:0 standard at2H values
  ungroup() %>% 
  summarize(F_0_at2H_mn = mean(at2H),
            F_0_at2H_se = sd(at2H))

# Add sF_0 to analyte data

peaks_mapped_with_threes_analytes <- peaks_mapped_with_threes_analytes %>% 
  mutate(F_0_at2H_se = 
           sF_0 %>% pull(F_0_at2H_se)
         )

```