### 2022-03-24 00_dD_label_measurements.Rmd
rename

---
title: "dD_label_measurements"
author: "Tristan Caro"
date: "3/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(tidyverse)
```

```{r}
vsmow <- 0.0001557643
permil_to_F <- function(dD) {
  frac <- (dD/1000 + 1) * vsmow / (1 + (dD/1000 + 1) * vsmow) * 1000000
  return(frac)
}

dD_df <- readxl::read_xlsx("data/Label_2H_Data/CU dD water data_Oct2021.xlsx",
                        sheet = 2,
                        skip = 1) %>% 
  filter(str_detect(`CH Vial Labes`, "CU-")) %>% 
  mutate(`F_label_ppm` = permil_to_F(as.numeric(dD_SMOW_corr))) %>% 
  mutate(inc_time_d = case_when(
    str_detect(Vial_label, "0x") ~ 0,
    str_detect(Vial_label, "0y") ~ 0,
    str_detect(Vial_label, "0z") ~ 0,
    str_detect(Vial_label, "3x") ~ 3,
    str_detect(Vial_label, "3y") ~ 3,
    str_detect(Vial_label, "3z") ~ 3,
    str_detect(Vial_label, "7x") ~ 7,
    str_detect(Vial_label, "7y") ~ 7,
    str_detect(Vial_label, "7z") ~ 7,
  )) %>% 
  mutate(soil = case_when(
    str_detect(Vial_label, "t") ~ "Tundra",
    str_detect(Vial_label, "c0") ~ "Conifer",
    str_detect(Vial_label, "g") ~ "Grassland",
    str_detect(Vial_label, "cr") ~ "Crust",
  ))

# Do we see changes in F_L over the course of the incubation? Over different terrains?
p_dD_df <- dD_df %>% 
  ggplot(aes(x = inc_time_d, y = F_label_ppm, color = soil)) +
  geom_smooth(method = "lm") +
  geom_point()

p_dD_df2 <- dD_df %>% 
  ggplot(aes(x = inc_time_d, y = as.numeric(dD_SMOW_corr), color = soil)) +
  geom_smooth(method = "lm") +
  geom_point()

dD_df %>% 
  group_by(soil) %>% 
  summarize(F_label_ppm_mn = mean(F_label_ppm),
            F_label_ppm_se = sd(F_label_ppm))
```

# Data export 
```{r}
write_csv(dD_df, file = "data/Label_2H_Data/dD_cleaned.csv")
```