### 2023-01-07 calibration_submodule/02_calibration.Rmd
calibration submodule commit

---
title: "Sample Data Analysis"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged # omit to disable paged data table output
    number_sections: yes # change to no for unnumbered sections
    toc: yes # change to no to disable table of contents
    toc_float: true # change to false to keep toc at the top
    toc_depth: 3 # change to specify which headings to include in toc
    code_folding: show # change to hide to hide code by default
editor_options:
  chunk_output_type: console # change to inline to show output inline
---

# Setup (*)

```{r setup, echo = TRUE, message=FALSE, warning=FALSE}
# load libraries
library(rlang)
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
library(purrr)
library(stringr)
library(readxl)
library(isoreader)
library(isoprocessor)

source("scripts/plotting_functions.R")
source("scripts/table_functions.R")
source("scripts/error_propagation.R")

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf", "postscript"),
  fig.keep = "all",
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.path = file.path("plots", paste0(gsub("\\.[Rr]md", "", knitr::current_input()), "_"))
)
```

```{r}
message("using isoreader version ", packageVersion("isoprocessor"))
message("using isoprocessor version ", packageVersion("isoprocessor"))
```

# Load Data (*)

```{r}
sample_peak_table <-
  readxl::read_excel("data/peak_table_raw_data.xlsx") %>%
  iso_make_units_implicit()
```

# Analyte Peaks Overview

```{r "analytes_overview", fig.width=8, fig.height=8, warning=FALSE}
# generate overview
sample_peak_table %>%
  filter(!is.na(compound) & compound != "H2") %>%
  iso_plot_data(
    x = area2, y = d2H, 
    color = type,
    label = analysis,
    size = 3,
    points = TRUE
  ) +
  scale_color_brewer(palette = "Dark2") + 
  scale_x_log10(breaks = c(1, 2, 3, 6, 10, 25, 50, 100, 200, 400)) +
  theme(panel.grid.minor = element_blank()) +
  facet_grid(type~., scales = "free_y") 
```

# Memory Calculation (*)

```{r}
# alternative memory calculation not used in the end but included for completeness
sum_memory <- function(area, delta) {
  stopifnot(length(area) == length(delta))
  mem <- 0
  mem_area <- 0
  for (i in seq_along(area)) {
    if (area[i] > mem_area) {
      # new area bigger --> takes over entire memory
      mem <- delta[i]
      mem_area <- area[i]
    } else {
      # new area smaller --> partially replace memory
      mem = ((mem_area - area[i]) * mem + area[i] * delta[i])/(mem_area)
    }
  }
  return(mem)
}

# calculate memory effect
memory_effects <- 
  sample_peak_table %>%
  # focus on analyte peaks
  filter(is.na(compound) | compound != "H2") %>%
  select(file_id, peak_id, compound, rt, area2, d2H_vs_H2) %>%
  # cross join all peaks
  { left_join(
    ., 
    select(., file_id, compound_i = compound, rt_i = rt, area2_i = area2, d2H_vs_H2_i = d2H_vs_H2) %>%
      iso_strip_units(),
    by = "file_id"
    ) } %>%
  # use prior peaks for memory calculation
  group_by(file_id, peak_id, rt) %>%
  filter(rt_i <= rt) %>%
  ungroup() %>%
  # summarize for each file and peak
  group_by(file_id, peak_id) %>%
  summarize(
    mem_d2H = 
      if_else(
        n() == 1, 0, 
        sum(d2H_vs_H2_i[rt_i < rt] * area2_i[rt_i < rt]) / sum(area2_i[rt_i < rt])),
    mem_d2H_sum = sum_memory(area2_i, d2H_vs_H2_i),
    .groups = "drop"
  )

sample_peak_table_with_memory <-
  sample_peak_table %>%
  left_join(memory_effects, by = c("file_id", "peak_id"))
```

# Add Standards' Isotope Data (*)

```{r}
standards <-
  tibble(
    standard = readxl::excel_sheets("data/standards.xlsx"),
    data = map(standard, ~readxl::read_excel("data/standards.xlsx", sheet = .x))
  ) %>%
  unnest(data) %>%
  filter(!is.na(true_d2H)) %>%
  select(type = standard, compound, true_d2H)

standards %>% iso_make_units_explicit() %>% knitr::kable(digits=2)

sample_peak_table_with_stds <-
  sample_peak_table_with_memory %>%
  iso_add_standards(stds = standards, match_by = c("type", "compound"))
```

# Check individual F8 standards

Applying univariate linear regression:

$$
\begin{aligned}
\delta_{peak/H2} &= \beta_0 + \beta_1 \cdot \delta_{analyte/VSMOW}
\end{aligned}
$$

```{r}
F8_w_calibs <- sample_peak_table_with_stds %>%
  filter(type == "F8") %>%
  iso_remove_problematic_peak_mappings() %>%
  iso_prepare_for_calibration(group_by = file_id) %>%
  iso_generate_calibration(
    # simple direct calibration model
    model = lm(d2H_vs_H2 ~ true_d2H)
  )

F8_w_calibs %>% iso_get_problematic_calibrations() -> problematic.calibs

F8_w_calibs <- F8_w_calibs %>% iso_remove_problematic_calibrations()

F8_w_calibs %>%
   iso_get_calibration_parameters(
    select_from_coefs = c(term, estimate, SE = std.error, signif),
    select_from_summary = c(fit_R2 = adj.r.squared, fit_RSD = sigma, residual_df = df.residual)
    ) %>%
  arrange(term) %>%
  head() %>%
  knitr::kable(digits=4)
```

# Check derived parameters

Look at the derived parameters to check consistency:

$$
\begin{aligned}

\delta_{H2/VSMOW} &= -\frac{\delta_{VSMOW/H2}}{\delta_{VSMOW/H2} + 1} = 
  \frac{ (1 - \beta_1) \cdot \delta_{mem/H2} - \beta_0}{(1 + \delta_{mem/H2} ) \cdot \beta_1} 
  \approx \frac{-\beta_0}{\beta_1} \\

\frac{A_{mem}}{A_{peak}} &= \frac{1 + \beta_0 - \beta_1}{1 + \delta_{mem/H2}} \approx 1 + \beta_0 - \beta_1 \\
& \rightarrow A_{mem} = \overline{A_{peak}} \cdot \left(1 + \beta_0 - \beta_1 \right)
\end{aligned}
$$

```{r "standards_parameters", fig.width=10, fig.height=8, warning=FALSE}
calc_d2H_vs_VSMOW <- rlang::expr(-value_b0/value_b1)
calc_percent_mem <- rlang::expr(1 + value_b0/1000 - value_b1)
calc_A_mem <- rlang::expr(mean_area * (1 + value_b0/1000 - value_b1))

stds_w_derived_params <- 
  F8_w_calibs %>%
  iso_get_calibration_data(select = c(analysis, mean_area_identified, mean_area_identified_sd, 
                                      injection_volume, folder, group, type)) %>%
  iso_get_calibration_parameters(
    select_from_coefs = c(term, estimate, std.error),
    select_from_summary = c()
  ) %>%
  mutate(
    term = case_when(term == "(Intercept)" ~ "b0", term == "true_d2H" ~ "b1"),
    mean_area = as.numeric(mean_area_identified),
    mean_area_sd = as.numeric(mean_area_identified_sd)
  ) %>%
  rename(value = estimate, se = std.error) %>%
  pivot_wider(names_from = term, values_from = c(value, se)) %>%
  mutate(
    `d2H H2/VSMOW [permil]` = !!calc_d2H_vs_VSMOW,
    `d2H H2/VSMOW [permil] error` = propagate_error(!!calc_d2H_vs_VSMOW, dvalue_b0 = se_b0, dvalue_b1 = se_b1),
    `Amem / Apeak` = !!calc_percent_mem,
    `Amem / Apeak error` = propagate_error(!!calc_percent_mem, dvalue_b0 = se_b0, dvalue_b1 = se_b1),
    `Amem [Vs]` = !!calc_A_mem,
    `Amem [Vs] error` = propagate_error(!!calc_A_mem, dmean_area = mean_area_sd, dvalue_b0 = se_b0, dvalue_b1 = se_b1)
  )

stds_w_derived_params %>%
  pivot_longer( cols = c(`d2H H2/VSMOW [permil]`, `Amem / Apeak`, `Amem [Vs]`)) %>%
  left_join(
    stds_w_derived_params %>% 
      pivot_longer( cols = c(`d2H H2/VSMOW [permil] error`, `Amem / Apeak error`, `Amem [Vs] error`)) %>%
      mutate(name = str_remove(name, " error")) %>%
      select(file_id, name, error = value),
    by = c("file_id", "name")
  ) %>%
  filter(mean_area > 1) %>%
  mutate(name = as_factor(name)) %>%
  iso_plot_data(
    x=mean_area_identified, y=value, y_error = error, label = analysis,
    size=mean_area_identified, panel = name ~ ., color = group,
    shape = type,
    points = TRUE,
    # indicate means of d2H and Amem
    geom_smooth(
      data = ~filter(.x, !str_detect(name, fixed("Apeak"))),
      method = "lm", mapping = aes(color = NULL), formula = y ~ 1,
      color = "black", linetype = 2
    ),
    # indicate 1/A dependence of Amem/Apeak
    geom_smooth(
      data = ~filter(.x, str_detect(name, fixed("Apeak"))),
      method = "lm", mapping = aes(color = NULL), formula = y ~ I(1/x),
      color = "black", linetype = 2
    )
  ) + 
  scale_x_log10(breaks = c(1, 2, 3, 7, 15, 30, 60)) + 
  labs(y = NULL)
```

# Generate Global Galibration (*)

$$
\begin{aligned}
\delta_{peak/H2} 
  &= \beta_0 + \beta_1 \cdot \delta_{analyte/VSMOW} + \beta_2 \cdot \frac{1}{A_{peak}} + \beta_3 \cdot \frac{  \delta_{analyte/VSMOW}}{A_{peak}} + \beta_4 \cdot \frac{\delta_{mem/H2}}{A_{peak}}
\end{aligned}
$$

```{r}
# analytes
sample_peak_table_for_calibration <- 
  sample_peak_table_with_stds %>%
  filter(!is.na(compound), compound != "H2") %>%
  mutate(row_id = row_number())

# global calib
global_calib <- 
  bind_rows(
    # standards
    sample_peak_table_for_calibration %>%
      filter(
        !str_detect(type, "sample"), is_std_peak, 
        mean_area_identified > iso_double_with_units(3, "Vs")
      ) %>%
      mutate(use_in_calib = TRUE),
    # samples
    sample_peak_table_for_calibration %>%
      filter(str_detect(type, "sample")) %>%
      mutate(use_in_calib = FALSE)
  ) %>%
  iso_prepare_for_calibration() %>%
  iso_generate_calibration(
    calibration = "d2H",
    model=c(
      b0_b1_only = lm(d2H_vs_H2 ~ true_d2H),
      plus_area_b2_b3 = lm(d2H_vs_H2 ~ true_d2H * I(1/area2)),
      plus_area_b2_b3_mem = lm(d2H_vs_H2 ~ true_d2H * I(1/area2) + mem_d2H : I(1/area2))
    ),
    use_in_calib = use_in_calib
  ) %>%
  iso_remove_problematic_calibrations()

# global calib for t0 (natural abundance)
global_calib_t0 <- 
  bind_rows(
    # standards
    sample_peak_table_for_calibration %>%
      filter(
        !str_detect(type, "sample"), is_std_peak, 
        mean_area_identified > iso_double_with_units(3, "Vs"),
        type == "F8"
      ) %>%
      mutate(use_in_calib = TRUE),
    # samples
    sample_peak_table_for_calibration %>%
      filter(type == "unlabeled sample") %>%
      mutate(use_in_calib = FALSE)
  ) %>%
  iso_prepare_for_calibration() %>%
  iso_generate_calibration(
    calibration = "d2H",
    model=c(
      b0_b1_only = lm(d2H_vs_H2 ~ true_d2H),
      plus_area_b2_b3 = lm(d2H_vs_H2 ~ true_d2H * I(1/area2)),
      plus_area_b2_b3_log = lm(d2H_vs_H2 ~ true_d2H * log(area2)),
      plus_area_b2_log = lm(d2H_vs_H2 ~ true_d2H + log(area2))
    ),
    use_in_calib = use_in_calib
  ) %>%
  iso_remove_problematic_calibrations()
```

# Global Calibration Parameters

```{r "global_calib_parameters", fig.width=8, fig.height=12}
global_calib %>%
  iso_get_calibration_parameters(
    select_from_coefs = c(),
    select_from_summary = 
      c(fit_R2=adj.r.squared, fit_RSD=sigma, residual_df=df.residual)
  ) %>%
  knitr::kable(digits=4)

global_calib_t0 %>%
  iso_get_calibration_parameters(
    select_from_coefs = c(),
    select_from_summary = 
      c(fit_R2=adj.r.squared, fit_RSD=sigma, residual_df=df.residual)
  ) %>%
  knitr::kable(digits=4)

global_calib %>% iso_plot_calibration_parameters()
global_calib_t0 %>% iso_plot_calibration_parameters() 
```

# Check derived parameters

$$
\begin{aligned}
\delta_{H2/VSMOW} &= -\frac{\delta_{VSMOW/H2}}{\delta_{VSMOW/H2} + 1} = \frac{-\beta_0}{\beta_0 + 1} =  \frac{1-\beta_1}{\beta_1} = -\frac{\beta_2}{\beta_2 - \beta_4} = -\frac{\beta_3 + \beta_4}{\beta_3} \\
A_{mem} &= \beta_4 = -\frac{\beta_2}{\beta_0} = - \frac{\beta_3}{\beta_1} = \beta_2 - \beta_3
\end{aligned}
$$

```{r}
calc_d2H_b0 <- rlang::expr(-val_b0 / (val_b0/1000 + 1))
calc_d2H_b1 <- rlang::expr(1000 * (1 - val_b1) / val_b1)
calc_d2H_b2_b4 <- rlang::expr(-val_b2 / (val_b2/1000 - val_b4))
calc_d2H_b3_b4 <- rlang::expr(- 1000 * (val_b3 + val_b4) / val_b3)
calc_Amem_b0_b2 <- rlang::expr(-val_b2 / val_b0)
calc_Amem_b1_b3 <- rlang::expr(-val_b3 / val_b1)
calc_Amem_b2_b3 <- rlang::expr(val_b2/1000 - val_b3)

calib_params <- 
  global_calib %>%
  iso_get_calibration_parameters(
    select_from_coefs = c(term, val = estimate, se = std.error),
    select_from_summary = c()
  ) %>%
  mutate(
    term = as_factor(term) %>%
      fct_recode(
        b0 = "(Intercept)", b1 = "true_d2H",
        b2 = "I(1/area2)", b3 = "true_d2H:I(1/area2)", 
        b4 = "I(1/area2):mem_d2H"
      )
  ) %>% 
  pivot_wider(values_from = c(val, se), names_from = term) %>%
  mutate(
    val_dH2_b0 = !!calc_d2H_b0,
    se_dH2_b0 = propagate_error(!!calc_d2H_b0, dval_b0 = se_b0),
    val_dH2_b1 = !! calc_d2H_b1,
    se_dH2_b1 = propagate_error(!!calc_d2H_b1, dval_b1 = se_b1),
    val_dH2_b2_b4 = !!calc_d2H_b2_b4,
    se_dH2_b2_b4 = propagate_error(!!calc_d2H_b2_b4, dval_b2 = se_b2, dval_b4 = se_b4),
    val_dH2_b3_b4 = !!calc_d2H_b3_b4,
    se_dH2_b3_b4 = propagate_error(!!calc_d2H_b3_b4, dval_b3 = se_b3, dval_b4 = se_b4),
    val_Amem_b4 = val_b4,
    se_Amem_b4 = se_b4,
    val_Amem_b0_b2 = !!calc_Amem_b0_b2,
    se_Amem_b0_b2 = propagate_error(!!calc_Amem_b0_b2, dval_b2 = se_b2, dval_b0 = se_b0),
    val_Amem_b1_b3 = !!calc_Amem_b1_b3,
    se_Amem_b1_b3 = propagate_error(!!calc_Amem_b1_b3, dval_b3 = se_b3, dval_b1 = se_b1),
    val_Amem_b2_b3 = !!calc_Amem_b2_b3,
    se_Amem_b2_b3 = propagate_error(!!calc_Amem_b2_b3, dval_b3 = se_b3, dval_b2 = se_b2)
  )

calib_params_table <- 
  left_join(
  calib_params %>%
    select(d2H_calib, starts_with("val")) %>%
    pivot_longer(cols = -d2H_calib, names_to = "var", 
                 names_transform = ~str_remove(.x, "val_")),
  calib_params %>%
    select(d2H_calib, starts_with("se")) %>%
    pivot_longer(cols = -d2H_calib, names_to = "var", values_to = "se", 
                 names_transform = ~str_remove(.x, "se_")),
  by = c("d2H_calib", "var")
) %>%
  mutate(
    formatted = format_with_decimals(value, error = se, decimals = find_signif_decimals(se, 1), include_plus = FALSE)
  ) %>%
  pivot_wider(id_cols = var, names_from = d2H_calib, values_from = formatted)

# results
calib_params_table

```

# Residuals

## Labeled samples

```{r "residuals", fig.width=11, fig.height=6, warning=FALSE}
global_calib %>%
  mutate(
    d2H_calib = as_factor(d2H_calib) %>%
      fct_recode(
        "$\\delta_{peak/H2} = \\beta_0 + \\beta_1 \\cdot \\delta_{analyte/VSMOW}$" = "b0_b1_only",
        "$... + \\beta_2 \\cdot \\frac{1}{A_{peak}} + \\beta_3 \\cdot \\frac{\\delta_{analyte/VSMOW}}{A_{peak}}$" = "plus_area_b2_b3",
        "$... + \\beta_4 \\cdot \\frac{\\delta_{mem/H2}}{A_{peak}}$" = "plus_area_b2_b3_mem"
      )
  ) %>%
  # plot residuals
  iso_plot_residuals(
    x = area2, # vs. area
    points = TRUE, size = 2, alpha = 0.5,
    color = true_d2H,
    panel = NULL,
    # residual smoothing fit
    trendlines = FALSE,
    value_ranges = FALSE,
    smooth = geom_smooth(
      mapping = aes(color = NULL, shape = NULL), method = "loess", formula = y ~ x
    )
  ) +
  # put on log x scale
  scale_x_log10() +
  scale_color_continuous(labels = function(x) sprintf("%+.0f\U2030", x)) +
  theme_figure(grid = FALSE) +
  theme(legend.position = c(0.93, 0.23), legend.title.align = 0.5) +
  labs(
    x = "mass 2 area (A) in Vs",
    y = "standard residuals in \U2030",
    color = expression(delta^2*H)
  ) +
  facet_grid(.~d2H_calib, labeller = latex_labeller)
```

## Unlabeled samples

```{r "residuals_t0", fig.width=9, fig.height=6, warning=FALSE}
global_calib_t0 %>%
  # plot residuals
  iso_plot_residuals(
    x = area2, # vs. area
    points = TRUE, size = 2, alpha = 0.5,
    shape = type,
    # residual smoothing fit
    trendlines = FALSE,
    smooth = geom_smooth(
      mapping = aes(color = NULL, shape = NULL), method = "loess", formula = y ~ x
    )
  ) +
  # put on log x scale
  scale_x_log10() +
  theme(legend.position = "right")
```

# Apply Global Calibrations (*)

## Labeled samples

```{r, warning=FALSE, cache=TRUE}
# applying model (also include without memory to figure out how much of a difference it makes)
global_calibs_applied <-
  global_calib %>%
  iso_apply_calibration(
    predict = true_d2H,
    calculate_error = TRUE
  )

# ranges
global_calibs_w_ranges <-
  global_calibs_applied %>%
  iso_evaluate_calibration_range(area2, true_d2H_pred)

global_calibs_w_ranges %>%
  iso_get_calibration_range() %>%
  knitr::kable(d=2)

calib_range_area2 <- c(
  iso_get_calibration_range(global_calibs_w_ranges)  %>% filter(term == "area2") %>% pull(min),
  to = iso_get_calibration_range(global_calibs_w_ranges)  %>% filter(term == "area2") %>% pull(max)
)
```

## Unlabeled samples

```{r, warning=FALSE, cache=TRUE}
# applying model (also include without area to check differences)
global_calibs_applied_t0 <-
  global_calib_t0 %>%
  filter(d2H_calib %in% c("b0_b1_only", "plus_area_b2_b3_log")) %>%
  iso_apply_calibration(
    predict = true_d2H,
    calculate_error = TRUE
  )

# ranges
global_calibs_w_ranges_t0 <-
  global_calibs_applied_t0 %>%
  iso_evaluate_calibration_range(area2, true_d2H_pred)

global_calibs_w_ranges_t0 %>%
  iso_get_calibration_range() %>%
  knitr::kable(d=2)

calib_range_area2_t0 <- c(
  iso_get_calibration_range(global_calibs_w_ranges_t0)  %>% filter(term == "area2") %>% pull(min),
  to = iso_get_calibration_range(global_calibs_w_ranges_t0)  %>% filter(term == "area2") %>% pull(max)
)
```

# Calibrated Data Table (*)

```{r}
# calibrated table - combine all the information
peak_table_calibrated <- global_calibs_w_ranges %>%
  iso_get_calibration_data() %>%
  filter(!is.na(true_d2H_pred)) %>%
  mutate(
    below_area_range = str_detect(d2H_in_range, fixed("<'area2'")),
    above_area_range = str_detect(d2H_in_range, fixed(">'area2'")),
    outside_d2H_range = str_detect(d2H_in_range, "true_d2H")
  ) %>% 
  select(-d2H_calib_params) %>%
  { 
    x <- filter(., d2H_calib == "b0_b1_only") %>%
      rename(
        calibrated_d2H_without_area = true_d2H_pred, 
        calibrated_d2H_without_area_se = true_d2H_pred_se
      )
    y <- filter(., d2H_calib == "plus_area_b2_b3") %>%
      select(
        peak_id, 
        calibrated_d2H_without_memory = true_d2H_pred, 
        calibrated_d2H_without_memory_se = true_d2H_pred_se
      )
    z <- filter(., d2H_calib == "plus_area_b2_b3_mem") %>%
      select(
        peak_id, 
        calibrated_d2H_with_memory_correction = true_d2H_pred, 
        calibrated_d2H_with_memory_correction_se = true_d2H_pred_se
      )
    left_join(x, y, by = "peak_id") %>%
      left_join(z, by = "peak_id")
  } %>%
  {
    x <- global_calibs_w_ranges_t0 %>%
      filter(d2H_calib == "b0_b1_only") %>%
      iso_get_calibration_data() %>%
      select(
        peak_id, 
        calibrated_d2H_unlabeled_no_area = true_d2H_pred, 
        calibrated_d2H_unlabeled_no_area_se = true_d2H_pred_se
      )
    y <- global_calibs_w_ranges_t0 %>%
      filter(d2H_calib == "plus_area_b2_b3_log") %>%
      iso_get_calibration_data() %>%
      select(
        peak_id, 
        calibrated_d2H_unlabeled = true_d2H_pred, 
        calibrated_d2H_unlabeled_se = true_d2H_pred_se
      )
    left_join(., x, by = "peak_id") %>%
      left_join(y, by = "peak_id")
  }
```

# Calibration Standards

```{r}
calibration_std_peak_table <-
  peak_table_calibrated %>%
  filter(d2H_in_calib )

# residuals + errors of the inverted calibration
deg_freedom_calib <- global_calib %>%
  filter(d2H_calib == "plus_area_b2_b3") %>%
  iso_get_calibration_summary() %>% pull(df.residual)

deg_freedom_calib_t0 <- global_calib_t0 %>%
  filter(d2H_calib == "plus_area_b2_b3_log") %>%
  iso_get_calibration_summary() %>% pull(df.residual)

# by compound
calibration_std_peak_table %>%
  group_by(compound, type) %>%
  summarize(
    calibrated_d2H_without_memory_sd = sd(calibrated_d2H_without_memory),
    calibrated_d2H_with_memory_correction_sd = sd(calibrated_d2H_with_memory_correction),
    .groups = "drop"
  )

# over all rmse
calibration_std_peak_table %>%
  summarize(
    n = n(), 
    # residual standard deviation (sometimes called incorrectly residual standard error)
    RDS_unlabeled = sqrt(sum((calibrated_d2H_unlabeled - true_d2H)^2, na.rm = TRUE) /
                 (deg_freedom_calib_t0)) %>% round(2),
    RSD_no_area = sqrt(sum((calibrated_d2H_without_area - true_d2H)^2, na.rm = TRUE) /
                 (deg_freedom_calib)) %>% round(2),
    RSD_no_mem = sqrt(sum((calibrated_d2H_without_memory - true_d2H)^2, na.rm = TRUE) /
                 (deg_freedom_calib)) %>% round(2),
    RSD_w_mem = sqrt(sum((calibrated_d2H_with_memory_correction - true_d2H)^2, na.rm = TRUE) /
                 (deg_freedom_calib)) %>% round(2),
    # root mean square error / root mean square deviation
    RMSE_unlabeled = sqrt(mean((calibrated_d2H_unlabeled - true_d2H)^2, na.rm = TRUE)) %>%
      round(2),
    RMSE_no_area = sqrt(mean((calibrated_d2H_without_area - true_d2H)^2, na.rm = TRUE)) %>%
      round(2),
    RMSE_no_mem = sqrt(mean((calibrated_d2H_without_memory - true_d2H)^2, na.rm = TRUE)) %>%
      round(2),
    RMSE_w_mem = sqrt(mean((calibrated_d2H_with_memory_correction - true_d2H)^2, na.rm = TRUE)) %>%
      round(2),
    .groups = "drop"
  ) %>%
  knitr::kable()
```

# Extraction Standards

```{r}
extraction_std_peak_table <- 
  peak_table_calibrated %>%
  filter(compound %in% c("21:0 (STD)", "23:0 (STD)"), !below_area_range)

extraction_std_peak_table %>%
  filter(!below_area_range, !above_area_range) %>%
  iso_plot_data(
    x = area2, y = c(calibrated_d2H_without_memory, calibrated_d2H_with_memory_correction), 
    color = compound,
    points = TRUE
  ) +
  scale_x_log10()

extraction_std_peak_table %>%
  filter(!below_area_range, !above_area_range) %>%
  group_by(compound) %>%
  summarize(
    calibrated_d2H_without_memory_sd = sd(calibrated_d2H_without_memory),
    calibrated_d2H_with_memory_correction_sd = sd(calibrated_d2H_with_memory_correction),
    .groups = "drop"
  )
```

# Sample Analytes

```{r}
# warning! 
# don't try to interpret peaks that are so small they are below the calibration area range,
# they are often overextrapolated
analytes_peak_table <-
  peak_table_calibrated %>%
  select(-use_in_calib, -is_std_peak, -individual, -injection_volume, -gc_method, -map_id, -peak_type, -ref_nr, -calib_peak, -is_ref, -d2H_in_calib, -d2H_calib, -true_d2H, -row_id, -d2H_resid) %>%
  filter(str_detect(type, "sample") & !str_detect(compound, "STD")) %>%
  mutate(
    calibrated_d2H_without_area = 
      ifelse(
        type == "unlabeled sample",
        calibrated_d2H_unlabeled_no_area,
        calibrated_d2H_without_area
      ),
    calibrated_d2H_without_area_se = 
      ifelse(
        type == "unlabeled sample",
        calibrated_d2H_unlabeled_no_area_se,
        calibrated_d2H_without_area_se
      ),
    calibrated_d2H_without_memory = 
      ifelse(
        type == "unlabeled sample",
        calibrated_d2H_unlabeled,
        calibrated_d2H_without_memory
      ),
    calibrated_d2H_without_memory_se = 
      ifelse(
        type == "unlabeled sample",
        calibrated_d2H_unlabeled_se,
        calibrated_d2H_without_memory_se
      ),
    note = ifelse(below_area_range, "careful interpreting these!", "")
  ) %>%
  select(
    -calibrated_d2H_unlabeled, -calibrated_d2H_unlabeled_se,
    -calibrated_d2H_unlabeled_no_area, calibrated_d2H_unlabeled_no_area_se
  ) %>%
  select(id1, type, everything()) %>%
  relocate(starts_with("calibrated_"), d2H_in_range, note, .after = last_col()) %>%
  iso_strip_units()

export_to_excel(analytes_peak_table, file = "output/samples_data.xlsx")
```