### 2024-02-20 00_methylation_correction.qmd
add methylation correction calculation

---
title: "Phthalic acid analysis"
date: "`r format(Sys.Date(), '%d %b %Y')`"
number-sections: true
number-offset: 0
toc: true
toc-depth: 3
df-print: tibble
embed-resources: true
output:
  html: 
    code-tools: true
    code-fold: show
    code-summary: "Show the code"
    toc-float: true
editor: source
editor_options: 
  chunk_output_type: console
---

Using `r R.version.string`, tidyverse version `r packageVersion("tidyverse")`, isoreader version `r packageVersion("isoreader")`, and isoprocessor version `r packageVersion("isoprocessor")`.

```{r}
#| label: setup
#| echo: true
#| message: false
#| warning: false

# load libraries
library(rlang)
library(tidyverse)
library(isoreader)
library(isoprocessor)

min_isoreader_version <- "1.3.9999"
min_isoprocesor_version <- "0.6.11"
if (packageVersion("isoreader") < min_isoreader_version) {
  sprintf(
    "min required isoreader version %s, install with devtools::install_github('isoverse/isoreader')", 
    min_isoreader_version) |> 
  stop(call. = FALSE)
}
if (packageVersion("isoprocessor") < min_isoprocesor_version) {
  sprintf(
    "min required isoprocessor version %s, install with devtools::install_github('isoverse/isoprocessor')", 
    min_isoprocesor_version) |> 
  stop(call. = FALSE)
}

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf"), fig.keep = "all",
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.path = file.path("plots", paste0(gsub("\\.[Rr](md|markdown)", "", knitr::current_input()), "_"))
)
```

# Calibration Prep

# Load data files

```{r}
# iso_files_raw <-
#   "data/PAME_data" |>
#   iso_read_continuous_flow(read_cache = TRUE, parallel = TRUE) |>
#   iso_save("data/PAME_data.cf.rds")

iso_files_raw <- 
  iso_read_continuous_flow("data/PAME_data.cf.rds") |>
  # remove files with read errors (usually aborted samples)
  iso_filter_files_with_problems() |>
  # process isodat peak table
  iso_set_peak_table_from_isodat_vendor_data_table() |>
  # convert units from mV to V for amplitudes and area
  iso_convert_peak_table_units(V = mV, Vs = mVs)
```

# Process File Information

```{r}
# parse file info
iso_files_w_file_info <- iso_files_raw |>
  # rename key file info columns
  iso_rename_file_info(id1 = `Identifier 1`, id2 = `Identifier 2`) |> 
  # process file information
  iso_mutate_file_info(
    # analysis as number
    analysis = as.integer(parse_number(Analysis)),
    # what is the type of each analysis?
    type = case_when(
      str_detect(id1, "[Zz]ero")      ~ "on_off",
      str_detect(id1, "H3")           ~ "H3_factor",
      str_detect(id1, "F8")           ~ "standard",
      str_detect(id1, "PAME")         ~ "pame",
      TRUE                            ~ "other"
    )
  ) |>
  # focus on the file info we want to use
  iso_select_file_info(analysis, file_datetime, id1, type)
```

# First look at chromatograms

```{r}
#| label: fig-chroms-first-look
#| warning: false
#| fig-width: 10
#| fig-height: 8
# look at specific files for peak mapping purposes

iso_files_w_file_info |> 
  iso_filter_files(analysis %in% c(12775, 12759)) |>
  iso_plot_continuous_flow_data(
    data = 2,
    #time_interval = c(1200, 1400),
    #time_interval = c(700, 1400),
    peak_label = iso_format(peak_nr, rt),
    panel = sprintf("%s (#%s)", id1, analysis),
    peak_label_options = list(size = 3)
  ) +
  theme(legend.position = "bottom")
```

# Peak Mapping

```{r}
# sample info and peak maps
peak_maps <- readxl::read_excel("data/PAME_peak_map.xlsx")

iso_files_w_mapped_peaks <- 
  iso_files_w_file_info |> 
  iso_filter_files(type %in% c("standard", "pame")) |>
  # map peaks
  iso_map_peaks(peak_maps, map_id = type)
iso_files <- iso_files_w_mapped_peaks
```

# Peak Table & Calculations

```{r}
#| label: all-peaks-calculations

# pull out the peak table and calculate the delta_peak vs. H2 ref
peak_table_all <- iso_files |>
  iso_get_peak_table(include_file_info = everything()) |>
  group_by(file_id) |>
  mutate(
    ref_ratio_at_rt = 
      lm(y ~ x, 
         data = tibble(
           x = rt[calib_peak == TRUE], 
           y = `r3/2`[calib_peak == TRUE])) |> 
      predict(newdata = tibble(x = rt)) |> 
      as.numeric(),
    d2H_vs_H2 = (`r3/2` / ref_ratio_at_rt - 1) * 1000
  ) |>
  ungroup()

# focus on analytes
peak_table <- peak_table_all |>
  filter(is.na(ref_nr)) |>
  iso_mutate_peak_table(
    group_by = file_id,
    rel_area = area2/sum(area2, na.rm=TRUE),
    mean_area = mean(area2, na.rm=TRUE),
    mean_area_identified = mean(area2[!is.na(compound)], na.rm=TRUE),
    sd_area_identified = sd(area2[!is.na(compound)], na.rm=TRUE),
    mean_amp_identified = mean(amp2[!is.na(compound)], na.rm=TRUE),
    contaminants = any(peak_type == "contaminant" & !is.na(d2H_vs_H2))
  )
```

# Add Standards' Isotope Data

```{r}
standards <- readxl::read_excel("data/standards.xlsx") |>
  mutate(
    type = "standard",
    true_d2H = iso_double_with_units(true_d2H, "permil")
  )

standards |> iso_make_units_explicit() |> knitr::kable(digits=2)

peak_table_with_stds <-
  peak_table |>
  iso_add_standards(stds = standards, match_by = c("type", "compound"))
```

# Generate Global Galibration

```{r}
# area_range
area_range <- c(10, 50)

global_calib_pame <- 
  peak_table_with_stds |>
  iso_remove_problematic_peak_mappings(remove_unidentified = FALSE, remove_ambiguous = FALSE) |>
  filter(type %in% c("standard", "pame")) |>
  iso_prepare_for_calibration() |>
  iso_generate_calibration(
    calibration = "d2H",
    model=c(calib = lm(d2H_vs_H2 ~ true_d2H)),
    # calibrate based on standards with all peaks
    use_in_calib = 
      type == "standard" & is_std_peak &
      # constrain area calibration range
      area2 >= iso_double_with_units(area_range[1], "Vs") &
      area2 <= iso_double_with_units(area_range[2], "Vs") 
  ) |>
  iso_remove_problematic_calibrations()
```

# Apply Calibration

For the narrow range where the pame peaks are, there's no point adding an area correction, instead just go with a linear correction.

```{r}
#| warning: false
global_calibs_pame_applied <-
  global_calib_pame |>
  iso_apply_calibration(
    predict = true_d2H,
    calculate_error = TRUE
  )

# ref tank (H2) isotope value estimate
beta_0 <- 
  global_calibs_pame_applied |> 
  iso_get_calibration_parameters(select_from_summary = c()) |> 
  filter(term == "(Intercept)")

d_H2_VSMOW.permil <- -beta_0$estimate/(1 + beta_0$estimate / 1000)
sprintf("Info: estimated d2H of the H2 ref tank (vs. VSMOW): %.1f permil", d_H2_VSMOW.permil)

# ranges
global_calibs_pame_w_ranges <-
  global_calibs_pame_applied |>
  iso_evaluate_calibration_range(area2, true_d2H_pred)

global_calibs_pame_w_ranges |>
  iso_get_calibration_range() |>
  knitr::kable(d=2)

# final peak table
peak_table_pame_final <- 
  global_calibs_pame_w_ranges |>
  iso_get_calibration_data() |>
  filter(!is.na(compound))
```

# Export

```{r "pame", fig.width=12, fig.height=9, warning = FALSE}
# phthala data
pame <- 
  peak_table_pame_final |>
  filter(peak_type == "analyte") |>
  iso_make_units_explicit(prefix = ".", suffix = "") |>
  rename(sample_id = id1) |>
  arrange(sample_id, compound, analysis)

pame_summary <- 
  pame |>
  summarize(
    pame_n = n(),
    pame_mean = mean(true_d2H_pred.permil),
    pame_sd = sd(true_d2H_pred.permil)
  )

# export
iso_export_data_to_excel(
  summary = pame_summary,
  data = pame, 
  filepath = "data/PAME_TAC_jan_2022.xlsx"
)
```

# Methanol Calculation

```{r}
pame_summary |>
  mutate(
    # known isotopic composition of the 4 non-exchangeable Hs of phthalic acid
    pa_d2H = -81.9,
    pa_d2H_err = 1.2,
    # mass balance calculation of methanol methyl group H d2H
    meoh_d2H = 10/6 * pame_mean - 4/6 * pa_d2H,
    meoh_d2H_err = sqrt((10/6 * pame_sd)^2 + (4/6 * pa_d2H_err)^2)
  ) |>
  knitr::kable(digits = 2)
```