### 2024-02-26 01C_Compare_PAME.qmd
comparison of non-corrected and corrected PAME data

---
title: "Compare PAME Correction Datasets"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

In this file, we compare the datasets which were generated with both un-corrected and corrected phthalic acid methyl ester (PAME) values.

```{r}
library(tidyverse)
```

## Import Datasets

```{r}
lipid_data_no_correction <- read_rds("cache/LH_SIP_data.RDS") |> 
  select(Analysis, sample_id, soil, inc_time_d, compound, area2, `2F_alk_mc`) |> 
  rename(`F2H_alk_not_corrected` = `2F_alk_mc`) |> 
  group_by(Analysis) |> 
  distinct(compound, .keep_all = TRUE)

lipid_data_corrected <- readxl::read_excel("data/LH_SIP_PAME_CORR.xlsx") |> 
  select(Analysis, sample_id, compound, `2F_alk_mc`) |> 
  rename(`F2H_alk_corrected` = `2F_alk_mc`) |> 
  group_by(Analysis) |> 
  distinct(compound, .keep_all = TRUE)
```

```{r}
lipid_data <- lipid_data_no_correction |> 
  left_join(lipid_data_corrected, by = c("Analysis", "sample_id", "compound")) |> 
  # filter non-negative F2H: some F2H are negative due to memory effect calculation
  # these negative F2H are filtered out in the growth rate script because
  # it is impossible to have negative growth
  filter(F2H_alk_not_corrected >= 0, F2H_alk_corrected >= 0) |> 
  mutate(diff_F2H = F2H_alk_not_corrected - F2H_alk_corrected) |> 
  mutate(F2H_alk_not_corrected_ppm = F2H_alk_not_corrected * 1e4,
         F2H_alk_corrected_ppm = F2H_alk_corrected * 1e4,
         diff_F2H_ppm = diff_F2H * 1e4)
```

# Compare

```{r}
p_line <- lipid_data |> 
  ggplot(
    aes(
      x = F2H_alk_not_corrected_ppm,
      y = F2H_alk_corrected_ppm
    )
  ) +
  geom_abline(color = "red") +
  stat_smooth(
    formula = "y~x", method = "lm", se = FALSE,
    color = "blue", linewidth = 0.5,
    fullrange = TRUE
  ) +
  #scale_x_continuous(limits = c(-3000, 3000)) +
  coord_cartesian(xlim = c(0, 2300), ylim = c(0, 2300)) +
  geom_point() +
  labs(
    x = "F2H (ppm) (before PAME correction)",
    y = "F2H (ppm) (after PAME correction)"
  ) +
  theme_bw() +
  theme(aspect.ratio = 1)
```

```{r}
mean_offset = lipid_data |> pull(diff_F2H) |> abs() |> mean()

p_histogram <- lipid_data |> 
  ggplot(
    aes(x = abs(diff_F2H) * 1e4)
  ) +
  geom_histogram(binwidth = 0.5) +
  geom_vline(xintercept = mean_offset * 1e4,
             color = "red", linetype = "dashed") +
  labs(
    x = "F2H difference (ppm)",
    y = "Count"
  ) +
  theme_bw() +
  theme(aspect.ratio = 1)

paste("The average offset between the data that was and was not properly corrected for the isotopic composition of the derivatization agent is", 
      round(mean_offset, 3), 
      "atom %; or ", 
      round(mean_offset * 1e4, 3), 
      "parts per million."
      )
```

```{r}
combined <- cowplot::plot_grid(p_line, p_histogram, nrow = 2)

cowplot::save_plot(plot = combined, base_height = 8, base_width = 6, filename = "fig_output/PAME_correction.pdf")
```