### 2024-07-25 99_BioCrust_Analysis.qmd
biocrust data analysis

---
title: "99: Biocrust Analysis"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# Setup

## Clear the environment

```{r}
rm(list=ls())
```

```{r setup, include = FALSE}
# global knitting options for code rendering
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf"),
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.keep = "all",
  fig.path = file.path("fig_output", paste0(gsub("\\.[Rr]md", "", knitr::current_input()), "_"))
)
```

## Load packages

```{r, message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)    # CRAN v1.3.1
library(forcats)      # CRAN v0.5.1 
library(isoreader)    # [::NA/NA] v1.3.0 # CRAN v1.3.0
library(isoprocessor) # [github::isoverse/isoprocessor] v0.6.7
library(isotopia)     # [github::isoverse/isotopia] v0.5.8 
library(readxl)       # CRAN v1.3.1 
library(ggsci)        # CRAN v2.9 
library(ggborderline) # [github::wurli/ggborderline] v0.1.0
library(ggrepel)      # CRAN v0.9.1 
library(latex2exp)    # CRAN v0.5.0
library(ggdist)       # CRAN v3.0.0 
library(ggsignif)     # CRAN v0.6.2
library(ggridges)     # CRAN v0.5.3
library(cowplot)      # CRAN v1.1.1
library(isotopia)
```

## Load sourced functions
These functions are found in the `libs/` directory.
```{r}
# Sourced Functions
source(file.path("libs", "visualization.R"))              # Visualization scripts
source(file.path("libs", "chromeleon.R"))                 # Chromeleon Reader
source(file.path("libs", "error_prop.R"))                 # Error Propagation
source(file.path("libs", "calculate_turnover.R"))         # Turnover Calculator
source(file.path("libs", "d2H_to_2F.R"))                  # Convert delta to frac abund
source(file.path("libs", "rename_FAs.R"))                 # Rename FAs to correctly ID'd FAs
var_to_str = function(v) {return(deparse(substitute(v)))} # Convert variable name to string
```

# Load data

## Load lipid d2H assimilation efficiency data

```{r}
d2H_assim <- read_xlsx("data/d2H_data_isotope_measurements.xlsx", sheet = 'lipids')
```

## Load F_L 2F Data

```{r}
F_L_2F <- read_csv("data/Label_2H_Data/dD_cleaned.csv") %>% 
  group_by(soil) %>% 
  summarize(F_label_ppm_mn = mean(F_label_ppm),
            F_label_ppm_se = sd(F_label_ppm)) %>% 
  mutate(F_label_mn = F_label_ppm_mn / 10000,
         F_label_se = F_label_ppm_se / 10000) %>% 
  mutate(terrain = case_when(
    soil == "Conifer" ~ "Gordon Gulch Conifer Forest",
    soil == "Tundra" ~ "Niwot Ridge Tundra",
    soil == "Grassland" ~ "Marshall Mesa Grassland"
  ))

write_csv(F_L_2F, file = "data/Label_2H_Data/dD_summarized.csv")
```

## Load Compound nC and nH Data sheet

```{r}
compound_nC_nH <- read_xlsx("data/compound_nC_nH.xlsx")
```

## Load IRMS Data

### Load Metadata and Problematic Run Data

```{r}
mtda <- read_csv(file.path("data", "sample_metadata_2021.csv"))
sample_names <- mtda$id1

problematic_runs <- read_xlsx(file.path("data", "problematic_runs.xlsx"))
problematic_analyses <- problematic_runs$Analysis
problematic_analyses_num <- parse_number(problematic_analyses)
```

### Load Peak Map

```{r}
# this information is often maintained in a csv or Excel file
peak_map <- 
  # initial peak map:
  # readxl::read_excel(file.path("data","IRMS", "peak_map_manual_general.xlsx"))
  # corrected peak map
  readxl::read_excel(file.path("data", "peak_maps_resolved.xlsx"))
```

### Read raw IRMS data files

```{r, warning=FALSE}
# Set file path(s) to data files, folders or rds collections 
# can be multiple folders or mix of folders and files
# Isoverse will iteratively search subfolders. Huzzah! ^_^
data_path <- file.path("data", "IRMS", "raw_data")
ref_ratio <- get_standard("2H") %>% as.numeric()

# read files
iso_files_raw <- 
  # path to data files
  data_path %>% 
  # read data files in parallel for fast read
  iso_read_continuous_flow() %>%
  # filter out files with read errors (e.g. from aborted analysis)
  iso_filter_files_with_problems()
```

### Process file info & peak table

```{r}
# process IRMS file information
iso_files <- iso_files_raw %>% 
  # rename key file info columns
  iso_rename_file_info(analysis = Analysis, id1 = `Identifier 1`, id2 = `Identifier 2`) %>% 
  # parse text info into numbers
  iso_parse_file_info(number = analysis) %>% 
  # process other file information that is specific to the naming conventions
  # of this particular sequence
  iso_mutate_file_info(
    # what is the type of each analysis?
    type = case_when(
      str_detect(id1, "[Zz]ero")      ~ "on_off",
      str_detect(id1, "H3")           ~ "H3_factor",
      str_detect(id1, "F8")           ~ "F8_std",
      str_detect(id1, "F9")           ~ "F9_std",
      TRUE                            ~ "sample"
    ),
    # what was the concentration? (assuming Preparation = concentration or volume)
    concentration = 
      ifelse(type == "std",  
             str_extract(Preparation, "[0-9.]+ ?ng( per |/)uL") %>% 
               parse_number() %>% iso_double_with_units("ng/uL"),
             NA),
    # what folder are the data files in? (assuming folder = sequence)
    folder = basename(dirname(file_path))
  ) %>% 
  # focus only on the relevant file info, discarding the rest
  iso_select_file_info(
    folder, analysis, file_datetime, id1, type, concentration
  ) %>% 
  # add in additional sample metadata (could be any info)
  # note: this would typically be stored in / read from a csv or excel file
  iso_add_file_info(
    read_csv(file.path("data", "sample_metadata_2021.csv")),
    join_by = "id1"
  )
```

```{r}
sample_files <- iso_filter_files(iso_files, type == "sample")
sample_files <- iso_filter_files(sample_files, str_detect(id1, "cr"))
```

# Load GC-FID data

```{r}
root <- file.path("data", "biocrust_FID_data", "2021-06-11_TAC_FAMEs_cr0f0x-m0f3zF3")
file_list = list.files(path = root)
file_list <- file_list %>% 
  str_subset("blank", negate = TRUE) %>% 
  str_subset("Blank", negate = TRUE) %>% 
  str_subset("GCMS-STD", negate = TRUE) %>% 
  str_subset("BLK", negate = TRUE) # remove procedural blanks
sample_names <- file_list %>% 
  str_replace("FID.....", "") %>% 
  str_replace("_100ul-nhex.xls", "") %>% 
  str_replace("_100ul.xls", "")

FID_all_data <-
  tibble(
    sample = sample_names,
    file = file_list,
    data = map(file.path(root, file), read_chromeleon_export),
    injection_details = map(data, "injection_details"),
    integration_results = map(data, "integration_results")
  ) %>% select(-data) |> 
  filter(str_detect(sample,))

# Unnest the data
FID_results <- FID_all_data %>% 
  select(-injection_details) %>% 
  unnest(integration_results) %>% 
  filter(nchar(`Peak Name`) > 0) %>% 
  # Specify chain length
  mutate(
    chain = `Peak Name` %>% 
      str_remove("\\d+-OH") %>% 
      parse_number() %>% 
      abs() %>% 
      {paste0("C", .)} %>% 
      factor()
      ) %>% 
  # Rename these columns so we can join the IRMS and FID data
  # Important to specify what data comes from FID!
  rename(
    "sample_id" = sample,
    "compound" = `Peak Name`,
    "rt_FID" = `Retention Time`,
    "area_FID" = `Area`,
    "height_FID" = `Height`,
    "rel_area_FID" = `Relative Area`,
    "rel_height_FID" = `Relative Height`,
    "amount_FID" = `Amount`
         ) %>% 
  # Make FID notation match IRMS notation
  mutate(
    compound = 
      case_when(
        str_detect(compound, "18:2 trans 9, 12") ~ "18:2 9, 12",
        str_detect(compound, "18:2 trans 9,12") ~ "18:2 9, 12",
        str_detect(compound, "18:2 9,12") ~ "18:2 9, 12",
        str_detect(compound, "16:1 cis-9") ~ "16:1 cis 9",
        str_detect(compound, "16:1 trans-9") ~ "16:1 trans 9",
        str_detect(compound, "18:1 cis-9") ~ "18:1 cis 9",
        str_detect(compound, "18:1 trans-9") ~ "18:1 trans 9",
        str_detect(compound, "21:0") ~ "21:0 (STD)",
        str_detect(compound, "23:0") ~ "23:0 (STD)",
        TRUE ~ compound
      )
  )
```

## Load memory-corrected IRMS data!

```{r}
LH_SIP_memcorr <- readxl::read_excel("data/samples_data_memory_corrected_20221228.xlsx") %>% 
# Trim it down
  select(-c(type, # or `standard` if using original mc dataset 
            #is_std_peak,
            folder, 
            file_datetime, 
            id2, 
            #individual,
            #injection_volume,
            #gc_method,
            #map_id, 
            gc_ramp, 
            group, 
            #peak_type,
            #ref_nr,
            #calib_peak, 
            peak_nr,
            #is_ref,
            rt_start,
            rt, 
            rt_end, 
            n_overlapping,
            n_matches,
            #use_in_calib,
            #d2H_in_calib, 
            #d2H_resid, 
            d2H_calib_points
            )
         ) %>% 
  filter(id1 %in% c("cr0f0x", "cr0f0y", "cr0f0z",
                     "cr0f3x", "cr0f3y", "cr0f3z",
                     "cr0f7x", "cr0f7y", "cr0f7z",
                     "cri0", "cri3", "cri7")) %>% 
  rename(sample_id = id1,
         Analysis = analysis) %>% 
  select(Analysis, sample_id, compound, area2, area3, note,
          below_area_range, 
          calibrated_d2H_without_area,
          calibrated_d2H_without_area_se,
          calibrated_d2H_without_memory, 
          calibrated_d2H_without_memory_se,
          calibrated_d2H_with_memory_correction, 
          calibrated_d2H_with_memory_correction_se) %>% 
  mutate(Analysis = as.numeric(Analysis)) %>% 
  # Convert Delta Values in Permil (vs. VSMOW) to fractional abundance (at%)
  mutate(
    # Without area correction:
    calibrated_at2H_without_area = d2H_to_2F(calibrated_d2H_without_area),
    calibrated_at2H_without_area_se = d2H_to_2F(calibrated_d2H_without_area_se),
    # With Area, without memory correction:
    calibrated_at2H_without_memory = d2H_to_2F(calibrated_d2H_without_memory),
    calibrated_at2H_without_memory_se = d2H_to_2F(calibrated_d2H_without_memory_se),
    # With memory correction:
    calibrated_at2H_with_memory_correction = d2H_to_2F(calibrated_d2H_with_memory_correction), 
    calibrated_at2H_with_memory_correction_se = d2H_to_2F(calibrated_d2H_with_memory_correction_se)
  )
```

# Join FID and memory-corrected IRMS data

```{r}
joined <- LH_SIP_memcorr |> 
  left_join(FID_results, join_by(sample_id, compound)) |> 
  select(sample_id, compound, calibrated_at2H_with_memory_correction, calibrated_at2H_with_memory_correction_se, area2, area_FID) |> 
  # add in incubation times
  mutate(
    inc_time_d = case_when(
      sample_id == "cri0" ~ 0,
      sample_id == "cri3" ~ 3,
      sample_id == "cri5" ~ 5,
      sample_id == "cri7" ~ 7,
      sample_id == "cr0f0x" ~ 0,
      sample_id == "cr0f0y" ~ 0,
      sample_id == "cr0f3x" ~ 3,
      sample_id == "cr0f3y" ~ 3,
      sample_id == "cr0f7x" ~ 7,
      sample_id == "cr0f7y" ~ 7,
      )
  ) |> 
  mutate(
    light_condition = case_when(
      str_detect(sample_id, "cri") ~ "Sun",
      str_detect(sample_id, "cr0") ~ "Dark"
    )
  ) |> 
  # average out replicates
  group_by(sample_id, compound, inc_time_d, light_condition) |> 
  summarize(
    calibrated_at2H_with_memory_correction = mean(calibrated_at2H_with_memory_correction),
    calibrated_at2H_with_memory_correction_se = mean(calibrated_at2H_with_memory_correction_se),
    area2 = mean(area2, na.rm = TRUE),
    area_FID = mean(area_FID, na.rm = TRUE)
  ) |> 
  ungroup()

```

Separate out the timepoint zero data, average it for an average t0 at2H value
```{r}
t0_data <- joined |> 
  filter(inc_time_d == 0) |> 
  summarize(t0_at2H = mean(calibrated_at2H_with_memory_correction),
            t0_at2H_se = mean(calibrated_at2H_with_memory_correction_se))
```

Import our label strength data

```{r}
FL_data <- read_csv("data/Label_2H_Data/dD_summarized.csv") |> 
  filter(soil == "Crust") |> 
  select(F_label_mn, F_label_se)
```

Import assim dataset
```{r}
d2H_assim <- readRDS("cache/d2H_assimilation.RDS")
```

Combine the datasets, create the abundance-weighted mean dataset

```{r}
pre_calculation_df <- joined |> 
  bind_cols(t0_data) |> 
  bind_cols(FL_data) |> 
  bind_cols(d2H_assim) |> 
  # apply area filter: 3Vs or greater!
  filter(area2 > 3)

# create abundance-weighted mean
weighted_df <- pre_calculation_df |> 
  ungroup() |> 
  group_by(sample_id, inc_time_d, light_condition) |> 
  mutate(
    relative_weight = area_FID / sum(area_FID, na.rm = TRUE)
  ) |> 
  summarize(
    at2H_weighted = 
      sum(calibrated_at2H_with_memory_correction * relative_weight, na.rm = TRUE) /
      sum(relative_weight, na.rm = TRUE),
    at2H_weighted_se = 
      sum(calibrated_at2H_with_memory_correction_se * relative_weight, na.rm = TRUE) /
      sum(relative_weight, na.rm = TRUE),
    at2H_t0_weighted = 
      sum(t0_at2H * relative_weight, na.rm = TRUE) /
      sum(relative_weight, na.rm = TRUE)
  ) |> 
  ungroup() |> 
  mutate(compound = "Weighted Mean")
```

Plot it:
```{r}
p_dF <- pre_calculation_df |> 
  ggplot(
    aes(
      x = inc_time_d,
      y = calibrated_at2H_with_memory_correction,
      color = compound
    )
  ) +
  geom_line(
    data = weighted_df,
    aes(y = at2H_weighted),
    color = "black",
    size = 1
  ) +
  geom_point(
    data = weighted_df,
    aes(y = at2H_weighted),
    color = "black",
    size = 3
  ) +
  geom_line(alpha = 0.5) +
  geom_point(size = 2) +
  scale_color_brewer(palette = "Set1") +
  facet_wrap(vars(light_condition)) +
  scale_x_continuous(breaks = c(0, 3, 7)) +
  labs(
    x = "Incubation Time (days)",
    y = latex2exp::TeX("$^2F$")
  ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    strip.background = element_blank()
  )

p_dF

cowplot::save_plot(
  filename = "biocrust_figs/p_dF.pdf",
  plot = p_dF, base_height = 4, base_width = 6
)
```

# Calculate turnover

```{r}

turnover <- pre_calculation_df |> 
  # CORRECT MC'd data so that it never is less than natural abundance at2H
  ungroup() |>                           # Evaluate each row individually
  mutate(
    # Calculate Turnover Rate in days from day 0 to 7 and 0 to 3
    u_d = calculate_turnover(
      a = a_mn,                      # Assimilation efficiency
      FT = calibrated_at2H_with_memory_correction,                   # 2F at time t (at%)
      F0 = t0_at2H,               # 2F at t0 (at%)
      FL = F_label_mn,               # 2F of label (at%)
      t = inc_time_d                     # change in time (days)
      ),

    # Calculate uncertainty in turnover rate
    su_d = calculate_sigma_mu(
      a = a_mn,                      # Assimilation efficiency
      F_T = calibrated_at2H_with_memory_correction,                   # 2F at time t (at%)
      F_0 = t0_at2H,               # 2F at t0 (at%)
      F_L = F_label_mn,               # 2F of label (at%)
      t = inc_time_d,                     # change in time (days)
      # Uncertainties:
      sF_0 = t0_at2H_se,       # uncertainty in IRMS measurement
      sF_L = F_label_se,        # uncertainty in label strength
      # uncertainty in FT due to memory-effect calculation:
      sF_T = calibrated_at2H_with_memory_correction_se, 
      sa = sa # uncertainty in assimilation efficiency
    )
  )

weighted_turnover <- weighted_df |> 
  ungroup() |> 
  mutate(
    u_d = calculate_turnover(
      a = 0.7069045,                      # Assimilation efficiency
      FT = at2H_weighted,                   # 2F at time t (at%)
      F0 = at2H_t0_weighted,               # 2F at t0 (at%)
      FL = 0.48385,               # 2F of label (at%)
      t = inc_time_d                     # change in time (days)
    ),
    su_d = calculate_sigma_mu(
      a = 0.7069045,                      # Assimilation efficiency
      F_T = at2H_weighted,                   # 2F at time t (at%)
      F_0 = at2H_t0_weighted,               # 2F at t0 (at%)
      F_L = 0.48385,               # 2F of label (at%)
      t = inc_time_d,                     # change in time (days)
      # Uncertainties:
      sF_0 = 0.01739246,       # uncertainty in IRMS measurement
      sF_L = 0.007610569,        # uncertainty in label strength
      # uncertainty in FT due to memory-effect calculation:
      sF_T = at2H_weighted_se, 
      sa = 0.1716424 # uncertainty in assimilation efficiency
    )
  )

turnover_combined <- turnover |> 
  bind_rows(weighted_turnover) |> 
  mutate(
    gen_d = log(2) / u_d
  )
```

Plot it:
```{r}
p_ud <- turnover_combined |> 
  mutate(
    is_wm = case_when(compound == "Weighted Mean" ~ TRUE)
  ) |> 
  filter(inc_time_d != 0) |> 
  mutate(
    compound = fct_reorder(compound, u_d)
  ) |> 
  ggplot(
    aes(
      y = compound,
      x = u_d,
      fill = is_wm
    )
  ) +
  geom_text(
    #data = function(df) df |> filter(is_wm),
    aes(label = paste(round(gen_d, 0))),
    x = 0.35,
    size = 3
  ) +
  geom_pointrange(
    aes(
      xmin = u_d - su_d,
      xmax = u_d + su_d
    ),
    shape = 21,
    color = "black"
  ) +
  facet_wrap(vars(light_condition, inc_time_d), ncol = 1) +
  coord_cartesian(xlim = c(0, 0.37)) +
  scale_fill_manual(values = c("#f56042", "black")) +
  labs(
    x = latex2exp::TeX("Biomass growth rate $(day^{-1})$"),
    y = ""
  ) +
  theme_classic() +
  theme(
    strip.background = element_blank(),
    legend.position = "None"
    )
p_ud

cowplot::save_plot(
  filename = "biocrust_figs/p_ud.pdf",
  p_ud, base_height = 6, base_width = 6
)
```

# Export
```{r}
writexl::write_xlsx(LH_SIP_memcorr, path = "biocrust_figs/biocrust_raw.xlsx")
writexl::write_xlsx(turnover_combined, path = "biocrust_figs/biocrust_turnover_data.xlsx")
```